---
title: "Piping"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
```


## Piping

### Setup

We're learning how to combine functions into pipelines, so let's load the `{dplyr}` package along with the `{palmerpenguins}` package to use the `penguins` data set.

```{r echo = TRUE, message = FALSE}
library(dplyr)
library(palmerpenguins)
head(penguins)
```


## `|>` vs. `%>%`

We will focus on using the base R pipe `|>`. But you should recognize the tidyverse pipe `%>%` if you see it in the wild (or in some of your old code). For most purposes, there are not major differences between the pipes. The key advantages of `|>` are that (1) it is in base R, so no additional packages are need and (2) it applies to base R functions as well as tidyverse function. `%>%` primarily works with tidyverse functions.

If you want to use the tidyverse pipe, you need to have the `{magrittr}` package loaded. If you load `{tidyverse}` or any tidyverse packages (e.g., `{dplyr}`, `{readr}`), the tidyverse pipe is automatically loaded.

A key difference between `|>` and `%>%` is the placeholder used when the data object is not the first argument for a function.

```{r summary-quiz}
quiz(
  question(
    "What does the base R `|>` pipe use as a placeholder?",
    answer("`_`", correct = TRUE),
    answer("`.`"),
    answer("`.x`"),
    answer("`?`"),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question(
    "What is wrong with the following code?\n```df |> select(df, col1:col4)```",
    answer("`df` should not be included again after the pipe", correct = TRUE),
    answer("You need to use a `%>%` pipe in the situation instead of a `|>` pipe"),
    answer("The placeholder is missing"),
    answer("The `select()` function has to be on a separate line from the pipe"),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```

## Using pipes

Pipes can be used to create whole chains of functions to wrangle data. In many cases, the order of the commands doesn't matter---unless it matters (ðŸ˜).  

### 

Please complete these commands in the order specified.

**Produce a data frame that (1) includes the species, sex, island, and body mass, (2) converts the species to a character, and (3) creates a new column called _body_mass_kg_ that converts body mass from grams to kilograms and replaces the _body_mass_g_ column using the `.keep` argument. Don't save the output to an object---just return the output to the console.**
```{r exercise1, exercise = TRUE}
penguins |> ...
```

```{r exercise1-check}
grade_this({
  if (identical(.result, penguins |>  
  select(species, sex, island, body_mass_g) |> 
  mutate(species = as.character(species),
         body_mass_kg = body_mass_g / 1000, .keep = "unused"))) {
  pass()
  }
  fail("Not quite there. Try again!")
})
```




### 

**Recreate the following data frame with as little code as possible.**
```{r echo = FALSE}
penguins |> 
  select(species, year, island, sex) |> 
  mutate(species_year = paste(species, year, sep = "_"), .after = year) |> 
  mutate(sex_num = if_else(sex == "female", 1, 0, NA))
```
```{r exercise2, exercise = TRUE}
penguins |> ...
```
```{r exercise2-check}
grade_this({
  if (identical(.result, penguins |> 
  select(species, year, island, sex) |> 
  mutate(species_year = paste(species, year, sep = "_"), .after = year) |> 
  mutate(sex_num = if_else(sex == "female", 1, 0, NA)))) {
  pass()
  }
  fail("Not quite there. Try again!")
})
```

### 

**Produce a data frame that (1) includes only the columns starting with "bill_", (2) converts all measurements from mm to cm (in one line; don't worry about changing the column names), and (3) creates a _bill_area_ column by multiplying the bill length and depth. **
```{r exercise3, exercise = TRUE}
penguins |> ...
```

```{r exercise3-solution}
penguins |>  
  select(starts_with("bill_")) |> 
  mutate(across(contains("_mm"), ~ . / 10),
         bill_area = bill_length_mm * bill_depth_mm)
```

```{r exercise3-code-check}
grade_code()
```

### 

### Wrap-up

Congratulations, you finished the tutorial!

To get credit for this assignment, replace my name with the first name that you submitted in the course introduction form in the code below and click *Run Code* to generate the text for you to submit to Canvas. 

```{r exit_function}
letter_nums <- letters[1:26]
generate_text <- function(x) {
  x <- tolower(x)
  offset <- 5
  if (!is.character(x)) stop("Oops! Did you surround your name with quotation marks?")
  if (x == "jeff") stop("Oops! You left my name in. Please type your name.")
  x <- stringr::str_sub(x, 1, 4)
  expanded <- stringr::str_split_fixed(x, "", 4)
  name_nums <- match(expanded, letter_nums)
  name_nums <- name_nums[!is.na(name_nums)]
  new_nums <- as.character(name_nums + offset)
  new_text <- paste("Now copy this string of numbers and submit them to Canvas:", stringr::str_c(new_nums, collapse = ""))
  return(new_text)
}
```

```{r exit_code, exercise = TRUE, exercise.setup = "exit_function"}
# replace my name below with your first name (surrounded by quotes)
first_name <- "Jeff"
generate_text(first_name)
```

### Assignment complete!

Great! Copy that code into Canvas, and you're all set for this tutorial.
