---
title: "Strings"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
options(tutorial.max.forked.procs = 10)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
set.seed(3)
```


## Introduction

### Character strings

First, let's load the `{stringr}` package.
```{r echo = TRUE}
library(stringr)
```


Character strings are a basic data type in R. When reading in data, typically any columns that have non-numeric data will be treated as character strings. 

As review, how do you coerce a vector of numeric data that you want to be characters? 

**1. Convert the vector `x` below to be a character string.**
```{r exercise1, exercise = TRUE}
x <- 1:10
x <- ...
```
 
```{r exercise1-solution}
x <- 1:10
x <- as.character(1:10)
```
 
```{r exercise1-check}
grade_result(
  pass_if(~identical(.result, x <- as.character(1:10)))
)
```

###

We assign strings by wrapping them in quotes. But what happens if we want to include quotes within strings?

```{r string-quotes}
question(
  'Which option(s) would generate this character string? \n
  _The central message of Buddhism is not "every man for himself"_',
  answer('`\'The central message of Buddhism is not "every man for himself"\'`', correct = TRUE),
  answer('`\"The central message of Buddhism is not \\"every man for himself\\"\"`', correct = TRUE),
  answer('`\"The central message of Buddhism is not \"every man for himself\"\"`'),
  answer('`\'The central message of Buddhism is not \\"every man for himself\\"\'`', correct = TRUE),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

## Working with strings

###

Several `{stringr}` functions allow you to work with string length, extraction, and case.

```{r string-length}
quiz(
  question(
    'What would be the output of the following command: `length(c("Avoid", "the", "green", "ones.", "They\'re", "not", "ripe", "yet."))`?',
    answer('`8`', correct = TRUE),
    answer('`5 3 5 5 7 3 4 4`'),
    answer('`"Avoid" "the" "green" "ones." "They\'re" "not" "ripe" "yet."`'),
    answer('`"Avoid the green ones. They\'re not ripe yet."`'),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question(
    'What would be the output of the following command: `str_length(c("Avoid", "the", "green", "ones.", "They\'re", "not", "ripe", "yet."))`?',
    answer('`8`'),
    answer('`5 3 5 5 7 3 4 4`', correct = TRUE),
    answer('`"Avoid" "the" "green" "ones." "They\'re" "not" "ripe" "yet."`'),
    answer('`"Avoid the green ones. They\'re not ripe yet."`'),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question(
    'What would be the output of the following command: `str_c(c("Avoid", "the", "green", "ones.", "They\'re", "not", "ripe", "yet."), collapse = " ")`?',
    answer('`8`'),
    answer('`5 3 5 5 7 3 4 4`'),
    answer('`"Avoid" "the" "green" "ones." "They\'re" "not" "ripe" "yet."`'),
    answer('`"Avoid the green ones. They\'re not ripe yet."`', correct = TRUE),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```

###

**2. Type the code that would extract the _first_ three characters from each element of the following vector:**
```{r exercise2, exercise = TRUE}
x <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
...
```
 
```{r exercise2-solution}
x <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
str_sub(x, 1, 3)
```
 
```{r exercise2-check}
grade_result(
  pass_if(~identical(.result, str_sub(x, 1, 3)))
)
```

###

**3. Type the code that would extract the _last_ three characters from each element of the following vector:**
```{r exercise2b, exercise = TRUE}
x <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
...
```
 
```{r exercise2b-solution}
x <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
str_sub(x, -3, -1)
```
 
```{r exercise2b-check}
grade_result(
  pass_if(~identical(.result, str_sub(x, -3, -1)))
)
```

###

**4. Type the code that converts the MLA formatted title _Beyond Good and Evil_ to APA formatted title _Beyond good and evil_.**
```{r exercise3, exercise = TRUE}
x <- "Beyond Good and Evil"
...
```
 
```{r exercise3-solution}
x <- "Beyond Good and Evil"
str_to_sentence(x)
```
 
```{r exercise3-check}
grade_result(
  pass_if(~identical(.result, str_to_sentence(x)))
)
```

## Combining strings with output

### Pasting and gluing

For the following exercises, use the appropriate function to create the sentence: `"Yes they do, Otto. They just don't understand it."`

**5. Produce this sentence use the object `name` and the `paste()` function.**
```{r exercise4, exercise = TRUE}
name <- "Otto"
paste(...)
```
 
```{r exercise4-solution}
name <- "Otto"
paste("Yes they do, ", name, ". They just don't understand it.", sep = "")
```
 
```{r exercise4-check}
grade_result(
  pass_if(~identical(.result, paste("Yes they do, ", name, ". They just don't understand it.", sep = "")))
)
```

###

**6. Now produce the sentence using the `paste0()` function.**
```{r exercise5, exercise = TRUE}
name <- "Otto"
paste0(...)
```
 
```{r exercise5-solution}
name <- "Otto"
paste0("Yes they do, ", name, ". They just don't understand it.")
```
 
```{r exercise5-check}
grade_result(
  pass_if(~identical(.result, paste0("Yes they do, ", name, ". They just don't understand it.")))
)
```

###

**7. Now produce the sentence using the `str_glue()` function with `name` as an argument.**
```{r exercise6, exercise = TRUE}
name <- "Otto"
str_glue(...)
```
 
```{r exercise6-solution}
name <- "Otto"
str_glue("Yes they do, {name}. They just don't understand it.", name = "Otto")
```
 
```{r exercise6-check}
grade_result(
  pass_if(~identical(.result, str_glue("Yes they do, {name}. They just don't understand it.", name = "Otto")))
)
```

### Wrap-up

Congratulations, you finished the tutorial!

To get credit for this assignment, replace my name with the first name that you submitted in the course introduction form in the code below and click *Run Code* to generate the text for you to submit to Canvas. 

```{r exit_function}
letter_nums <- letters[1:26]
generate_text <- function(x) {
  x <- tolower(x)
  offset <- 6
  if (!is.character(x)) stop("Oops! Did you surround your name with quotation marks?")
  if (x == "jeff") stop("Oops! You left my name in. Please type your name.")
  x <- stringr::str_sub(x, 1, 4)
  expanded <- stringr::str_split_fixed(x, "", 4)
  name_nums <- match(expanded, letter_nums)
  name_nums <- name_nums[!is.na(name_nums)]
  new_nums <- as.character(name_nums + offset)
  new_text <- paste("Now copy this string of numbers and submit them to Canvas:", stringr::str_c(new_nums, collapse = ""))
  return(new_text)
}
```

```{r exit_code, exercise = TRUE, exercise.setup = "exit_function"}
# replace my name below with your first name (surrounded by quotes)
first_name <- "Jeff"
generate_text(first_name)
```

### Assignment complete!

Great! Copy that code into Canvas, and you're all set for this tutorial.
