---
title: "Plotting x-y data: associations"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(palmerpenguins)
library(Hmisc)
options(tutorial.max.forked.procs = 10)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
```


## Introduction

This exercise covers visualizing x-y data, focusing on associations.

### Set-up

Load `{tidyverse}` and `{palmerpenguins}`.

```{r echo = TRUE, message = FALSE}
library(tidyverse)
library(palmerpenguins)
```

## Scatterplots

###

Scatterplots plot paired continuous data. Here we see (again) the penguin bill length plotted against bill depth.

```{r scatter, echo = TRUE, warning = FALSE}
penguins %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm)) +
  geom_point()
```

### Adding regression lines

Just as a refresher, you can add regression lines with `geom_smooth(method = "lm")`.

```{r scatter-lm, echo = TRUE, warning = FALSE, message = FALSE}
penguins %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm)) +
  geom_point() +
  geom_smooth(method = "lm")
```

###

**Now add separate regression lines for each species.**

```{r scatter-lm2, warning = FALSE, message = FALSE}
penguins %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, color = species)) +
  geom_point() +
  geom_smooth(method = "lm")
```

```{r exercise1, exercise = TRUE, warning = FALSE, message = FALSE}
penguins %>% 
  ...
```

```{r exercise1-solution}
penguins %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, color = species)) +
  geom_point() +
  geom_smooth(method = "lm")
```

### Adding diagonal lines

If you are plotting two variables that are on the same scale (e.g., pre/post, measure 1 vs. measure 2), you may want to add a diagonal line in which _x=y_. For this, we can use the `geom_abline()` function and specify the `intercept` as 0 and the `slope` as 1. Unfortunately, there are no paired variables on the same scale in the `penguins` dataset, so we'll draw a line where _2.5x=y_. That is, let's imagine that we want to plot the line showing where bill length is 2.5 times bill depth. 

**Plot the bill depth and length data including a line with slope is 2.5 and y-intercept is 0.**

```{r scatter-line, warning = FALSE}
penguins %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 2.5)
```

```{r exercise2, exercise = TRUE, warning = FALSE}
penguins %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm)) +
  ...
```

```{r exercise2-solution}
penguins %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 2.5)
```

## Correlation plots

Since `{ggplot2}` does not have a built in way to easily generate a correlation plot, we're going to use base R's `pairs()` function. You can also use the `{GGally}` function `ggpairs()`.

We'll plot the correlations between all of the size measurements in `penguins`.

```{r pairs, echo = TRUE}
penguins %>% 
  select(contains("_mm") | contains("_g")) %>% 
  pairs(upper.panel = NULL)
```



## Bubble charts

### Size as a third variable

Bubble charts vary the size of points to represent information and can be used in two ways. First, they can be used with continuous data in scatterplots to represent a third variable with the size of the data points. Here we plot bill depth and length of Adelie penguins as the x and y axes. Then we map the point size to body mass to generate the bubble chart.

```{r bubble1, echo=TRUE, warning=FALSE}
penguins %>% 
  filter(species == "Adelie") %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, size = body_mass_g)) +
  geom_point()
```

###

It is difficult to see all of the individual points.

**Replicate the above plot, but make the points 50% transparent.**

```{r bubble2, warning=FALSE}
penguins %>% 
  filter(species == "Adelie") %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, size = body_mass_g)) +
  geom_point(alpha = 0.5)
```

```{r exercise3, exercise = TRUE, warning = FALSE}
penguins %>% 
  filter(species == "Adelie") %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm)) +
  ...
```

```{r exercise3-solution}
penguins %>% 
  filter(species == "Adelie") %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, size = body_mass_g)) +
  geom_point(alpha = 0.5)
```

### Size as count

Bubble charts can also show the number of observations that fall within each x-y position. These can be used for both continuous and discrete data. For these plots, we use the `geom_count()` function.

```{r bubble3, echo=TRUE}
penguins %>%
  ggplot(aes(x = island, y = year)) +
  geom_count()
```

### 

In this case, year is treated as a numeric, which creates strange y-axis values. We'll learn other ways to deal with this later, but for now, let's convert year to a factor and replot.

**Coerce year to a factor and plot the above bubble chart.**

```{r bubble4}
penguins %>%
  mutate(year = factor(year)) %>% 
  ggplot(aes(x = island, y = year)) +
  geom_count()
```

```{r exercise4, exercise = TRUE, warning = FALSE}
penguins %>%
  ...
```

```{r exercise4-solution}
penguins %>%
  mutate(year = factor(year)) %>% 
  ggplot(aes(x = island, y = year)) +
  geom_count()
```



### Wrap-up

Congratulations, you finished the tutorial!

To get credit for this assignment, replace my name with the first name that you submitted in the course introduction form in the code below and click *Run Code* to generate the text for you to submit to Canvas. 

```{r exit_function}
letter_nums <- letters[1:26]
generate_text <- function(x) {
  x <- tolower(x)
  offset <- 13
  if (!is.character(x)) stop("Oops! Did you surround your name with quotation marks?")
  if (x == "jeff") stop("Oops! You left my name in. Please type your name.")
  x <- stringr::str_sub(x, 1, 4)
  expanded <- stringr::str_split_fixed(x, "", 4)
  name_nums <- match(expanded, letter_nums)
  name_nums <- name_nums[!is.na(name_nums)]
  new_nums <- as.character(name_nums + offset)
  new_text <- paste("Now copy this string of numbers and submit them to Canvas:", stringr::str_c(new_nums, collapse = ""))
  return(new_text)
}
```

```{r exit_code, exercise = TRUE, exercise.setup = "exit_function"}
# replace my name below with your first name (surrounded by quotes)
first_name <- "Jeff"
generate_text(first_name)
```

### Assignment complete!

Great! Copy that code into Canvas, and you're all set for this tutorial.
