---
title: "Group challenges"
output: html_document
date: "2023-03-06"
execute:
  freeze: auto
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Group data wrangling challenge

Using the `penguins` data set from the {palmerpenguins} package, recreate this data frame.
```{r}
library(tidyverse)
library(palmerpenguins)
```
```{r echo = FALSE}
penguins |> 
  filter(species != "Gentoo") |> 
  group_by(species, island, sex) |> 
  summarise(mean_body_mass = mean(body_mass_g, na.rm = TRUE)) |> 
  drop_na(sex) |> 
  pivot_wider(id_cols = c(species, island), names_from = sex, values_from = mean_body_mass)
```

## Group plotting challenge

plot1 <- penguins |>
  filter(species == "Adelie") |>
  drop_na(sex) |>
  mutate(Sex = fct_recode(sex, "Female" = "female",
                          "Male" = "male")) |>
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm, color = Sex)) +
  geom_point(size = 2.8, alpha = 0.55) +
  geom_smooth(method = "lm", se = F, color = "gray50") + 
  scale_x_continuous(limits = c(31, 46), n.breaks = 15) +
  scale_y_continuous(limits = c(16, 21), n.breaks = 8,
                     expand = c(0, 1.5)) +
  annotate(geom = "segment", x = 33.5, xend = 34.35, 
           y = 21.5, yend = 21.2, color = "black") +
  annotate(geom = "text", label = "Penguin 42", x = 34, y = 21.8) +
  labs(y = "Bill depth (mm)", x = "Bill length (mm)", title = "Bill size by sex") +
  theme_classic() +
  theme(legend.position = c(0.85, 0.2),
        legend.background = element_rect(fill = "transparent", color = "transparent"))

plot2 <- penguins |>
  filter(species == "Adelie") |>
  drop_na(sex, island) |>
  mutate(Sex = fct_recode(sex, "Female" = "female",
                          "Male" = "male"),
         Island = island) |>
  ggplot(aes(x = Sex, y = flipper_length_mm, 
             group = Island, shape = Island, color = Island)) +
  stat_summary(position = position_dodge(width = 0.12)) +
  stat_summary(fun = mean, geom = "line", position = position_dodge(width = 0.12)) +
  labs(y = "Flipper length (mm)", 
       title = "Mean and standard error of flipper length by sex and island") +
  theme_classic() + 
  theme(legend.position = c(0.85, 0.25))

plot3 <- penguins |>
  filter(species == "Adelie") |>
  drop_na(sex, island, year) |>
  mutate(Sex = fct_recode(sex, "Female" = "female",
                          "Male" = "male"),
         Year = as.factor(year)) |>
  ggplot(aes(x = Year, y = flipper_length_mm, color = Year)) +
  geom_jitter(width = .2, alpha = 0.6) +
  stat_summary(fun.data = mean_cl_normal, color = "black") +
  scale_y_continuous(limits = c(170, 210), n.breaks = 5) +
  geom_text(stat = "count", aes(label = paste0("N = ", after_stat(count))), y = 170,
            color = "black") +
  facet_wrap(vars(island)) +
  labs(y = "Flipper length (mm)", 
       title = "Mean and 95% confidence interval of flipper length by island and year") +
  scale_color_manual(values = c("#0072B2", "#009E73", "#E69F00")) +
  theme_classic() +
  theme(legend.position = "none")

(plot1 + plot2) / plot3 + 
  plot_annotation(title = "Adelie penguin bill and flipper size",
                  caption = "Source {palmerpenguins} data set",
                  tag_levels = "A")
