---
title: "Numbers"
author: "Jeff Stevens"
date: "2023-03-06"
date-format: iso
execute:
  echo: true
  freeze: true
format:
  revealjs: 
    theme: custom.scss
    slide-number: true
    code-line-numbers: false
    highlight-style: github
    code-overflow: wrap
    footer: "[DPaViR 2023](https://jeffreyrstevens.quarto.pub/dpavir)"
    code-link: true
---


# Introduction


## The problem

```{r echo = FALSE}
library(tidyverse)
nrows <- 12
set.seed(311)
data1 <- tibble(val1 = runif(nrows), val2 = runif(nrows), val3 = runif(nrows, 0.001, 0.01))#, val4 = paste0("$", format(runif(nrows, 10, 100), digits = 4)))
data2 <- data1 |> 
  # mutate(min = sec / 60) |> 
  mutate(percent_val1 = round(val1 * 100, 2), .keep = "unused", .before = 1) |> 
  mutate(log_val2 = log(val2), .keep = "unused", .after = percent_val1) |> 
  mutate(val3 = format(val3, scientific = TRUE, digits = 2)) #|> 
  # mutate(val4 = parse_number(val4))
```

What's different between these data sets? 

What is needed to create `data2` from `data1`?

:::: {.columns}

::: {.column width="45%"}

```{r}
data1
```
:::

::: {.column width="55%"}

```{r}
data2
```
:::

::::


## Set-up

```{r}
library(tidyverse)
library(nycflights13)
```


## Types of numbers

Doubles are floating point numbers

::: {.callout-note}
**Floating point number:** a number without a fixed number of digits after the decimal point
:::

Floating point numbers ≈ scientific notation


## Types of numbers

Computer memory is limited, so you cannot store numbers with infinite precision and floating points are stored imprecisely

```{r}
sqrt(2) ^ 2 == 2
1 / 49 * 49 == 1
```


# Comparing and counting


## Comparing numbers

:::: {.columns}

::: {.column width="50%"}

```{r}
x <- 1
y <- 1.00000000000001
z <- 1.001
```

::: {.fragment}
```{r}
x == y
all.equal(x, y)
all.equal(x, z)
all.equal(x, z, tolerance = 1e-2)
```
:::
:::

::: {.column width="50%"}

::: {.fragment}
```{r}
sqrt(2) ^ 2 == 2
all.equal(sqrt(2) ^ 2, 2)
dplyr::near(sqrt(2) ^ 2, 2)
```
:::

:::
::::

<!-- ## Parsing numbers -->

<!-- Sometimes, numbers come in as character string, occasionally with weird formatting. -->

<!-- To parse numbers that come in with scientific notation, use `readr::parse_double()`: -->
<!-- ```{r} -->
<!-- (num_char <- "1e3") -->
<!-- typeof(num_char) -->
<!-- parse_double(num_char) -->
<!-- ``` -->


<!-- ## Parsing numbers -->

<!-- To parse numbers that come in with monetary units or percent signs, use `readr::parse_number()`: -->
<!-- ```{r} -->
<!-- (num_char <- c("$1,234", "USD 3,513", "59%")) -->
<!-- typeof(num_char) -->
<!-- parse_number(num_char) -->
<!-- ``` -->


## Counts

As a reminder, we've already seen how to use `dplyr::count()`
```{r}
count(flights, carrier)
```


## Counts

:::: {.columns}

::: {.column width="50%"}

We can also automatically sort by count.

```{r}
count(flights, carrier, sort = TRUE)
```
:::

::: {.column width="50%"}
And sum up totals instead of just count

```{r}
count(flights, carrier, wt = distance)
```
:::

::::


## Counts

:::: {.columns}

::: {.column width="50%"}

Remember `n()` counts inside a `summarise()`

```{r}
flights |> 
  group_by(carrier) |> 
  summarise(n = n())
```
:::

::: {.column width="50%"}

::: {.fragment}

`n_distinct()` counts instances within a group
```{r}
flights |> 
  group_by(dest) |> 
  summarise(carriers = n_distinct(carrier))
```
:::
:::

::::


## Counting NAs

To count `NA`s, you can `sum()` up `TRUE` responses to `is.na()`

```{r}
flights |> 
  group_by(dest) |> 
  summarize(n_cancelled = sum(is.na(dep_time)))
```


## Counting NAs

This trick can be used for any logical vector

```{r}
sum(flights$month == 1)
nrow(filter(flights, month == 1))
```


# Transforming numbers

## Operations

Mathematical operators are recycled across all elements in a vector

```{r}
0:10 * 5
```

::: {.fragment}
You can operate with vectors > 1, but the larger vector must be a multiple of the smaller vector
```{r warning = TRUE}
0:10 * c(5, 6)
```
```{r}
0:11 * c(5, 6)
```
:::


## Mathematical transformations

:::: {.columns}

::: {.column width="30%"}

* `sqrt()`
* `log()`
* `log10()`
* `log2()`
* `sin()`
* `asin()`
:::

::: {.column width="70%"}

```{r}
sqrt(seq(0, 100, 10))
log(runif(10))
asin(sqrt(runif(10)))
```
:::
::::

## Rounding

Control significant digits with `round()`

```{r}
round(123.456, 2)  # two digits
round(123.456, 1)  # one digit
round(123.456) # whole number
round(123.456, -1) # round to nearest ten
round(123.456, -2) # round to nearest hundred
```

<!-- ## Rounding -->

<!-- But weirdness for 0.5s with even digits -->

<!-- ```{r} -->
<!-- round(c(1.5, 2.5)) -->
<!-- ``` -->

<!-- To avoid, use `ceiling()` -->

<!-- ```{r} -->
<!-- ceiling(c(1.5, 2.5)) -->
<!-- ``` -->


# Formatting numbers


## Formatting

When numbers get too big, too small,  or need other formatting, use `format()`

:::: {.columns}

::: {.column width="50%"}

```{r}
(x <- 0.0020)
format(x, scientific = TRUE)
format(x, nsmall = 4)
```
:::

::: {.column width="50%"}

::: {.fragment}


```{r}
(y <- 12345678.9)
format(y, scientific = TRUE, 
       digits = 3)
format(y, big.mark = ",")
```
:::
:::
::::


## Cutting numbers into ranges

If you need to bin numbers into ranges, use `cut()`

```{r}
set.seed(1)
(x <- runif(12, min = 0, max = 100))
cut(x, breaks = c(0, 33, 66, 100))
```
::: {.fragment}
```{r}
cut(x, breaks = c(0, 33, 66, 100), labels = c("Low", "Medium", "High"))
```
:::


<!-- ## Offsets -->

<!-- You many need to work with numbers after (`dplyr::lag()`) or before (`dplyr::lead()`) the current element -->

<!-- ```{r} -->
<!-- (x <- c(2, 5, 11, 11, 19, 35)) -->
<!-- lag(x) -->
<!-- x - lag(x) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- lead(x) -->
<!-- x - lead(x) -->
<!-- ``` -->


<!-- # Summarizing numbers -->

<!-- ## Summary functions -->

<!-- :::: {.columns} -->

<!-- ::: {.column width="50%"} -->

<!-- * `mean()` -->
<!-- * `median()` -->
<!-- * `min()` -->
<!-- * `max()` -->
<!-- * `sd()` -->
<!-- ::: -->

<!-- ::: {.column width="50%"} -->

<!-- * `IQR()` -->

<!-- ```{r} -->
<!-- IQR(x) -->
<!-- ``` -->

<!-- * `quantile()` -->

<!-- ```{r} -->
<!-- quantile(x) -->
<!-- ``` -->
<!-- ::: -->
<!-- :::: -->


## Solving the problem

:::: {.columns}

::: {.column width="45%"}

```{r}
data1
```
:::

::: {.column width="55%"}

```{r}
data2
```
:::

::::


## Let's code!

[Numbers](../code/19_numbers.html){target="_blank"} [[Rmd](../code/19_numbers.Rmd){target="_blank"}]
