---
title: "Functions"
author: "Jeff Stevens"
date: "2023-03-27"
date-format: iso
execute:
  echo: true
  freeze: auto
format:
  revealjs: 
    theme: custom.scss
    slide-number: true
    code-line-numbers: false
    highlight-style: github
    code-overflow: wrap
    footer: "[DPaViR 2025](https://jeffreyrstevens.quarto.pub/dpavir)"
    code-link: true
---


# Introduction


<!-- ## The problem -->

<!-- ```{r echo = FALSE} -->
<!-- library(tidyverse) -->
<!-- nrows <- 12 -->
<!-- set.seed(12) -->
<!-- data1 <- tibble(test_date = as_date(sample(as_date("2023-01-01"):as_date("2023-01-31"), nrows, replace = TRUE)), -->
<!--                     birth_date = as.character(as_date(sample(as_date("1950-01-01"):as_date("2005-01-31"), nrows, replace = TRUE))) -->
<!--                     ) |>  -->
<!--   arrange(test_date) -->
<!-- data1$birth_date[4] <- "Jan 9, 1960" -->
<!-- data2 <- data1 |>  -->
<!--   mutate(birth_date = as_date(birth_date, format = c("%Y-%m-%d", "%b %d, %Y"))) |>  -->
<!--   filter(birth_date < "1990-01-01") |>  -->
<!--   mutate(age_at_testing = test_date - birth_date, -->
<!--          day_of_birth = wday(birth_date, label = TRUE, abbr = FALSE)) -->
<!-- ``` -->

<!-- What's different between these data sets?  -->

<!-- :::: {.columns} -->

<!-- ::: {.column width="35%"} -->

<!-- ::: {style="font-size:90%"} -->
<!-- ```{r} -->
<!-- data1 -->
<!-- ``` -->
<!-- ::: -->
<!-- ::: -->

<!-- ::: {.column width="65%"} -->

<!-- ::: {style="font-size:90%"} -->
<!-- ```{r} -->
<!-- data2 -->
<!-- ``` -->
<!-- ::: -->
<!-- ::: -->

<!-- :::: -->


## Set-up

```{r}
library(tidyverse)
library(palmerpenguins)
```


## Why write your own functions?


# Creating functions

## Creating functions

```{r}
functionname <- function(argument1, argument2) {
  # Function contents
}
```


## Creating functions

```{r}
mymean <- function(x) {
  sum(x) / length(x)
}
```

:::{.fragment}
```{r}
mymean(mtcars$mpg)
mean(mtcars$mpg)
```
:::


## Multi-line functions
```{r}
mymean2 <- function(x) {
  mysum <- sum(x)
  mysum_divided <- mysum / length(x)
}
```

::: {.fragment}
```{r}
mymean2(mtcars$mpg)
```

Why no output?
:::

::: {.fragment}
```{r}
(mymean_obj <- mymean2(mtcars$mpg))
```
:::


## Returning output

Don't assign last step to object
```{r}
mymean3 <- function(x) {
  mysum <- sum(x)
  mysum / length(x)
}

mymean3(mtcars$mpg)
```

## Returning output

Or use `return()`
```{r}
mymean4 <- function(x) {
  mysum <- sum(x)
  mysum_divided <- mysum / length(x)
  return(mysum_divided)
}

mymean4(mtcars$mpg)
```


## Return multiple output values

`list()`

```{r}
mymean5 <- function(x) {
  mysum <- sum(x)
  mysum_divided <- mysum / length(x)
  list(sum = mysum, mean = mysum_divided)
}

mymean5(mtcars$mpg)
```


## Save intermediate objects

```{r}
mymean6 <- function(x) {
  mysum <<- sum(x)
  mysum / length(x)
}

mymean6(mtcars$mpg)
```

::: {.callout-note}
This is fine when testing out a function but probably not great practice 'in production'. Why?
:::


## Print messages to console

```{r}
mymean7 <- function(x) {
  mysum <- sum(x)
  print(mysum)
  mysum / length(x)
}

mymean7(mtcars$mpg)
```


## Print messages to console

```{r}
mymean8 <- function(x) {
  mysum <- sum(x)
  print(paste0("The sum is: ", mysum))
  mysum / length(x)
}

mymean8(mtcars$mpg)
```


## Print messages to console

```{r message = TRUE}
mymean9 <- function(x) {
  mysum <- sum(x)
  message(paste0("The sum is: ", mysum))
  mysum / length(x)
}

mymean9(mtcars$mpg)
```

::: {.callout-note}
Check out [`{cli}`](https://cli.r-lib.org/) package for powerful messaging.
:::

# Arguments

## Arguments

```{r}
multiplier <- function(x, constant) {
  x * constant
}
```

::: {.fragment}
```{r}
multiplier(x = 7, constant = 3)
multiplier(x = 1:10, constant = 3)
```

```{r eval = FALSE}
multiplier(x = 1:10)
```
```
Error in multiplier(x = 1:10) : 
argument "constant" is missing, with no default
```
:::


## Argument default values

```{r}
multiplier2 <- function(x, constant = 3) {
  x * constant
}
```

```{r}
multiplier2(x = 1:10)
multiplier2(x = 1:10, constant = 5)
```

## Let's write a function!

Here's the formula that reverse codes scale values

`Value`~`max`~ `+` `Value`~`min`~ ` -Score`

Write a function that calculates the reversed code score. What arguments do you need?


# Conditional execution

## Conditional execution

```{r}
head(penguins[, 1:5])
```

```{r}
mymean6(penguins$bill_length_mm)
```



## Conditional execution

```{r}
head(penguins[, 1:5])
```

```{r}
mymean10 <- function(x) {
  sum(x, na.rm = TRUE) / sum(!is.na(x))
}
```

```{r}
mymean10(penguins$bill_length_mm)
```


## Conditional execution

But if you want the user to control whether `NA` is ignored
```{r}
mymean11 <- function(x, ignore_na = TRUE) {
  if (ignore_na) {
    sum(x, na.rm = TRUE) / sum(!is.na(x))
  } else {
    sum(x) / length(x)
  }
}
```

```{r}
mymean11(penguins$bill_length_mm)
mymean11(penguins$bill_length_mm, ignore_na = FALSE)
```

## Multiple conditions

Use `else if`

```{r}
age_cutoffs <- function(x) {
  if(x <= 1.5) {
    "puppy"
  } else if (x <= 3) {
    "adolescent"
  } else if (x <= 10) {
    "adult"
  } else {
    "senior"
  }
}
```

::: {.fragment}
```{r}
age_cutoffs(1)
age_cutoffs(2)
age_cutoffs(5)
```
:::


## Stopping based on conditionals

```{r}
age_cutoffs2 <- function(x) {
  if(x <= 1.5) {
    "puppy"
  } else if (x <= 3) {
    "adolescent"
  } else if (x <= 10) {
    "adult"
  } else if (x <= 20) {
    "senior"
  } else {
    stop("Age exceeded 20.")
  }
}
```

::: {.fragment}
```{r}
age_cutoffs2(15)
```
```{r eval = FALSE}
age_cutoffs2(22)
```
```
Error in age_cutoffs2(22) : Age exceeded 20.
```
:::


## Multiple conditions

Use `switch()`

```{r}
central_tend <- function(x, type) {
  switch(type,
         mean = mean(x),
         median = median(x),
         trimmed = mean(x, trim = .1))
}
```

::: {.fragment}
```{r}
vector <- rcauchy(100)
central_tend(x = vector, type = "mean")
central_tend(x = vector, type = "median")
central_tend(x = vector, type = "trimmed")
```
:::


# Data frame functions

## Creating functions in tidyverse

```{r}
mean_species_bill <- function(df) {
  df |> 
    group_by(species) |> 
    summarise(mean(bill_length_mm, na.rm = TRUE))
}

mean_species_bill(penguins)
```


## Creating functions in tidyverse

What if we want user to input grouping and response variable?

```{r}
grouped_mean <- function(df, group_var, mean_var) {
  df |> 
    group_by(group_var) |> 
    summarise(mean(mean_var, na.rm = TRUE))
}
```

::: {.fragment}
```{r eval = FALSE}
penguins |> 
  grouped_mean(group_var = species, mean_var = bill_length_mm)
```
```
Error in `group_by()`:
! Must group by variables found in `.data`.
✖ Column `group_var` is not found.
Run `rlang::last_trace()` to see where the error occurred.
```
:::

## Embracing

Embrace variables in `{{ }}`

```{r}
grouped_mean2 <- function(df, group_var, mean_var) {
  df |> 
    group_by({{ group_var }}) |> 
    summarize(mean({{ mean_var }}, na.rm = TRUE))
}
```

```{r}
penguins |> 
  grouped_mean2(group_var = species, mean_var = bill_length_mm)
```


## Let's code!

[Functions](../code/24_functions.html){target="_blank"} [[Rmd](../code/24_functions.Rmd){target="_blank"}]
