---
title             : "Tables"

urlcolor: "blue"

documentclass     : "apa6"
classoption       : "doc"
output            : papaja::apa6_pdf
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

First, load `{tidyverse}` and `{papaja}` and import the data.

```{r}
library(tidyverse)
library(papaja)
hai <- read_csv("https://decisionslab.unl.edu/data/thayer_stevens_2020_data1.csv") |>
  select(experiment, participant, condition, gender,
         panas_pre = panas_pre_pos, panas_post = panas_post_pos)
hai_long <- hai |>
  pivot_longer(contains("panas"), names_to = "prepost", values_to = "panas") |>
  mutate(condition = fct_recode(condition, "Control" = "control", "HAI" = "hai"),
         prepost = fct_recode(prepost, "Pre" = "panas_pre", "Post" = "panas_post"),
         prepost = fct_relevel(prepost, c("Pre", "Post")))
```

Now let's build a data frame that will be our table.

```{r}
(condition_prepost_means <- hai_long |>
  group_by(condition, prepost) |>
  summarise(across(starts_with("panas"), ~ mean(.x, na.rm = TRUE))))
```

\newpage

# Tables by `{knitr}`

The [`{knitr}`](https://yihui.org/knitr/) package uses the `kable()` function to format tables.

```{r}
library(knitr)
kable(condition_prepost_means)
```

\newpage
## Column and row names

You can control column names and row names with `col.names` and `row.names`.

```{r}
kable(condition_prepost_means, 
      col.names = c("Condition", "Prepost", "Mean PANAS"))
```

\newpage
## Column alignment

By default, character columns are left aligned and numeric columns are right aligned. You can set alignment manually with the `align` argument with `l` = left, `c` = center, and `r` = right. You can just pass a character string with a series of those letters.

```{r }
kable(condition_prepost_means,
      col.names = c("Condition", "Prepost", "Mean PANAS"), 
      align = "rcl")
```

\newpage
## Digit rounding

Round the digits for all numeric data columns with `digits` argument.

```{r }
kable(condition_prepost_means,
      col.names = c("Condition", "Prepost", "Mean PANAS"), 
      digits = 2)
```

If you want different digits for different columns, you can pass a vector to the `digits` argument.
```{r }
condition_prepost_means |> 
  pivot_wider(id_cols = condition, names_from = prepost, values_from = panas) |> 
  kable(digits = c(0, 2, 3))
```


\newpage
## Table titles

Add a title to the table with the `caption` argument. The good news is that we can cross-reference easily (Table \@ref(tab:title-table)). The bad news is that with captions, tables in PDFs are automatically placed at the top of the page. We'll see how to fix this later.

```{r title-table}
kable(condition_prepost_means,
      col.names = c("Condition", "Prepost", "Mean PANAS"), 
      caption = "PANAS scores by condition and prepost")
```

<!-- \newpage -->
<!-- ## LaTeX format -->

<!-- You can also set the `format` to be `"latex"`, which will output the LaTeX code for the table. You only need to do this if you want to use some LaTeX options to format the table (e.g., turn the table to landscape orientation, split the table across pages, etc.). Here we used the LaTeX-specific `booktabs` argument to make the table look nicer. -->
<!-- ```{r} -->
<!-- kable(condition_prepost_means, format = "latex", booktabs = TRUE, -->
<!--       col.names = c("Condition", "Prepost", "Mean PANAS")) -->
<!-- ``` -->
<!-- While using the LaTeX formatting allows for powerful control over your table, the downside is that it can only output to PDFs (not Word or HTML). Though some publishers allow you to submit LaTeX files as the editable version of a manuscript, some only allow Word documents. So you'll either need to find another way to export the table to Word or recreate the table in Word. -->

\newpage

# Supplementing kable with `{kableExtra}`

The `kable()` function is intentionally simple to use and therefore does not have a lot of additional functionality. The [`{kableExtra}`](https://haozhu233.github.io/kableExtra/) package supplements the `kable()` functionality with additional formatting options by adding additional functions after the `kable()` function call using the `|>` pipe (a bit like how `ggplot()` works). Check out [Create Awesome LaTeX Table with knitr::kable and kableExtra](https://haozhu233.github.io/kableExtra/awesome_table_in_pdf.pdf).

```{r}
# install.packages("kableExtra")
library(kableExtra)
```

## General styling

The `kable_styling()` function formats a number of things such as font size, table width, and table alignment. I'll also add  `latex_options = "hold_position"` to keep the table in the text. Otherwise, it puts it at the top of the page.

```{r}
kable(condition_prepost_means,
      caption = "PANAS scores by condition and prepost",
      col.names = c("Condition", "Prepost", "Mean PANAS"), 
      booktabs = TRUE) |>
  kable_styling(font_size = 18, latex_options = "hold_position")
```


\newpage
## Labels spanning rows

If you want to label groups of rows, use `pack_rows()`. Let's get rid of the `condition` column and label the conditions explicitly.

```{r}
condition_prepost_means2 <- condition_prepost_means |>
  ungroup(condition) |>
  select(-condition)
```

```{r}
kable(condition_prepost_means2, 
      booktabs = TRUE)
```

```{r}
kable(condition_prepost_means2, 
      booktabs = TRUE,
      col.names = c("", "Mean PANAS")) |>
  pack_rows("Control", start_row = 1, end_row = 2) |>
  pack_rows("HAI", start_row = 3, end_row = 4)
```

Notice that we removed the first column name with `""`.

\newpage
## Labels spanning columns

You can labels groups of columns with  the `add_header_above()` function. Let's rearrange the data into wide format to illustrate this.

```{r}
(wide_means <- condition_prepost_means |>
   unite(cond_prepost, condition:prepost) |>
   pivot_wider(names_from = cond_prepost, values_from = panas))
```


```{r}
kable(wide_means, booktabs = TRUE)
```

Now that the data are in wide format, we can add the column names by repeating _Pre_ and _Post_ then add the headers.

```{r}
kable(wide_means, 
      booktabs = TRUE, 
      col.names = rep(c("Pre", "Post"), 2),
      digits = 2) |>
  add_header_above(c("Control" = 2, "HAI" = 2))
```
Maybe we need a column stating this is Mean PANAS.

```{r}
wide_means2 <- wide_means |> 
  mutate(label = "Mean PANAS", .before = 1)
kable(wide_means2, 
      booktabs = TRUE, 
      col.names = c("", rep(c("Pre", "Post"), 2)),
      digits = 2) |>
  add_header_above(c("", "Control" = 2, "HAI" = 2))
```

\newpage
## Table footnotes

Add table notes with the `footnote()` function.

```{r}
kable(condition_prepost_means, 
      booktabs = TRUE,
      caption = "PANAS scores by condition and prepost",
      col.names = c("Condition*", "Prepost", "Mean PANAS")) |>
  kable_styling(latex_options = "hold_position") |>
  footnote(general = "Source: Thayer & Stevens (2021)",
           symbol = "73 control participants, 72 HAI participants",
           footnote_as_chunk = TRUE)
```

\newpage
## Landscape

Rotate wide tables with `landscape()` function.

```{r}
kable(condition_prepost_means, 
      booktabs = TRUE,
      caption = "PANAS scores by condition and prepost",
      col.names = c("Condition", "Prepost", "Mean PANAS")) |>
  kable_styling(latex_options = "hold_position") |>
  footnote(general = "Source: Thayer & Stevens (2021)",
           footnote_as_chunk = TRUE) |>
  landscape()
```


# Tables by `{papaja}`

The `{papaja}` package uses the `kable()` function to format tables in APA format with the `apa_table()` function. You can use many of the same arguments that are available in the `kable()` function. You can control where the table is placed (`h`ere, `t`op, `b`ottom) with the `placement` argument. You can add a general footnote with the `note` argument.

```{r}
apa_table(condition_prepost_means,
          caption = "PANAS scores by condition and prepost",
          col.names = c("Condition", "Prepost", "Mean PANAS"),
          placement = "h",
          note = "Source: Thayer & Stevens (2021)")
```

Notice the alignment is different, with everything left aligned. Let's right align the means.
```{r}
apa_table(condition_prepost_means, 
          caption = "PANAS scores by condition and prepost",
          col.names = c("Condition", "Prepost", "Mean PANAS"),
          align = c("l", "l", "r"),
          placement = "h",
          note = "Source: Thayer & Stevens (2021)")
```

\newpage

You can rotate to landscape orientation with the `landscape = TRUE` argument.

```{r}
apa_table(condition_prepost_means, 
          caption = "PANAS scores by condition and prepost",
          col.names = c("Condition", "Prepost", "Mean PANAS"),
          align = c("l", "l", "r"),
          placement = "h",
          note = "Source: Thayer & Stevens (2021)",
          landscape = TRUE)
```


# APA-formatted statistics by `{papaja}`

`{papaja}` also includes  `apa_print()`, which extracts statistical values in APA format.


## Linear regression

```{r}
hai_lm <- lm(panas_post ~ condition * gender, data = hai)
summary(hai_lm)
apa_print(hai_lm)
```

```{r lm-table}
apa_table(apa_print(hai_lm)$table,
          caption = "Linear regression results",
          placement = "h")
```

\newpage
Let's clean up those predictor names.

```{r}
hai_lm_table <- apa_print(hai_lm)$table
hai_lm_table <- hai_lm_table |>
  mutate(term = str_replace(term, "Conditionhai", "Condition"),
         term = str_replace(term, "GenderMale", "Gender"))
apa_table(hai_lm_table,
          caption = "Linear regression results",
          placement = "h")
```

**How could we name the first column _Predictor_ instead of _term_?**

\newpage

# Other table packages

* [`{gt}`](https://gt.rstudio.com/index.html) RStudio's grammar of tables (logically like ggplot2)
* [`{flextable}`](https://davidgohel.github.io/flextable/articles/overview.html) Good Word output but a bit tricky to work with
* [`{huxtable}`](https://hughjonesd.github.io/huxtable/) Very flexible but tricky to work with
