---
title: "Data types"
author: "Jeffrey R. Stevens"
date: "2023-02-03"
output: html_document
execute:
  freeze: auto
---

## Creating objects

### First, let's create some objects: 
`aa <- 3; bb <- 3L; cc <- "3"; dd <- "TRUE"; ee <- TRUE; ff <- "NA"; gg <- NA`
```{r}
aa <- 3
bb <- 3L
cc <- "3"
dd <- "TRUE"
ee <- TRUE
ff <- "NA"
gg <- NA
```

## Checking data types

### Guess what data type each object is then check it.
```{r}
typeof(aa)
typeof(bb)
typeof(cc)
typeof(dd)
typeof(ee)
typeof(ff)
typeof(gg)
```

### How do we test if `aa` is an integer?
```{r}
is.integer(aa)
```

### What will `is.logical(dd)` return?
```{r}
is.logical(dd)
```


### How do we test if `ff` and `gg` are `NA`?
```{r}
is.na(ff)
is.na(gg)
```

## Checking if objects are the same

### Are `aa` and `bb` the same? How do we test this?
```{r}
aa
bb
aa == bb
```

### What about `aa` and `cc`?
```{r}
aa
cc
aa == cc
```

### A safer comparison tool is `identical()`. Test if `aa` and `bb` are identical. Then try `aa` and `cc`.
```{r}
identical(aa, bb)
identical(aa, cc)
```

### Now see if `aa` is identical to `3` and if `bb` is identical to `3L`.
```{r}
identical(aa, 3)
identical(bb, 3L)
```

