---
title: "Data structures"
author: "Jeffrey R. Stevens"
date: "2025-02-05"
output: html_document
execute:
  freeze: auto
---

<!-- ## In-class coding -->

<!-- Make a sequence from 0 to 100 in steps of 10. -->
<!-- ```{r} -->
<!-- seq(0, 100, 10) -->
<!-- ``` -->

<!-- Create a repetition of “yes” and “no” with 10 instance of each, alternating between the two. Then make one with 10 "yes" and then 10 "no". -->
<!-- ```{r} -->
<!-- rep(c("yes", "no"), times = 10) -->
<!-- rep(c("yes", "no"), each = 10) -->
<!-- ``` -->


<!-- Add the argument `n = 10` to `head(mtcars)`. What does this do? -->
<!-- ```{r} -->
<!-- head(mtcars, n =  10) -->
<!-- ``` -->


<!-- ## Extra coding practice -->

### Vectors

Create a vector called  `dog_names` with the values _Bella_, _Daisy_, and _Max_.
```{r}
dog_names <- c("Bella", "Daisy", "Max")
```

Create a vector called `sex` with the values _Female_, _Male_, and _Male_.
```{r}
sex <- c("Female", "Female", "Male")
```

Use the index operator to print to console only _Daisy_ and _Max_ from `dog_names`.
```{r}
dog_names[2:3]
```

Replace the _Daisy_ entry with _Luna_ and print `dog_names` to console.
```{r}
dog_names[2] <- "Luna"
dog_names
```

Test whether the name _Sophie_ is contained in the vector `dog_names`.
```{r}
"Sophie" %in% dog_names
```


### Lists

Copy/paste and run this code: `(mylist <- list(a = 1:4, b = c(4, 3, 8, 5), c = LETTERS[10:15], d = c("yes", "yes")))`

```{r}
(mylist <- list(a = 1:4, b = c(4, 3, 8, 5), c = LETTERS[10:15], d = c("yes", "yes")))
```

Check the data types for each list element individually.
```{r}
typeof(mylist$a)
typeof(mylist$b)
typeof(mylist$c)
typeof(mylist$d)
```

Check the data types for each list element with one command.
```{r}
str(mylist)
```

Combine list elements `a` and `b` into a single vector.
```{r}
c(mylist$a, mylist$b)
```

### Data frames

Create a data frame called `mydf` with three columns: `x`, `y`, and `z` and five rows. For `x` assign any five numbers, for `y` assign any five character strings, and for `z` assign any five logical values.
```{r}
(mydf <- data.frame(x = sample(1:10, 5, replace = TRUE), y = sample(letters, 5), z = sample(c(TRUE, FALSE), 5, replace = TRUE)))
```

Create a data frame called `dogs` that combines the `dog_names` and `sex` vectors and print to console.
```{r}
(dogs <- data.frame(dog_names, sex))
```

Print to console just Luna's row.
```{r}
dogs[2, ]
```

Print to console the number of rows in `dogs`.
```{r}
nrow(dogs)
```

