---
title: "Pivoting data"
author: "Jeffrey R. Stevens"
date: "2023-02-24"
output: html_document
execute:
  freeze: auto
---

For these exercises, we'll use a new clean version of the dog breed traits data set. 

1. Import data from <https://jeffreyrstevens.quarto.pub/dpavir/data/dog_breed_traits_clean.csv> and assign to `traits`.

```{r}
library(tidyverse)
traits <- read_csv("https://jeffreyrstevens.quarto.pub/dpavir/data/dog_breed_traits_clean.csv")
```

2. Create `traits2` where we delete the coat columns, so we only have breed and ratings data.

```{r}
traits2 <- traits |> 
  select(-contains("coat"))
```

3. Is `traits2` tidy?

4. Is `traits2` in wide or long format?

5. Reshape `traits2` so that all of the ratings scores are in a single column called _rating_ with a column labeling what kind of rating it is called _scale_. Assign this to `traits3`.

```{r}
traits3 <- traits2 |> 
  pivot_longer(affectionate:grooming, names_to = "scale", values_to = "rating")
```

6. How would we check if `traits3` has the expected number of rows?

7. Create `traits4` by removing the rows with affectionate, children, and other_dogs as values of _scale_.

```{r}
traits4 <- traits3 |> 
  filter(!scale %in% c("affectionate", "children", "other_dogs"))
```

8. Spread out the data into wide format with separate columns for the shedding and grooming data, then create a new column _diff_ that subtracts grooming from shedding ratings.

```{r}
traits4 |>
  pivot_wider(id_cols = breed, names_from = scale, values_from = rating) |> 
  mutate(diff = shedding - grooming)
```

