---
title: "Separating and uniting data"
author: "Jeffrey R. Stevens"
date: "2023-02-27"
output: html_document
execute:
  freeze: auto
---

For these exercises, we'll use a new clean version of the dog breed traits data set. 

1. Load tidyverse, import data from <https://jeffreyrstevens.github.io/dpavir2025/data/dog_breed_traits_clean.csv>, and assign it to `traits`.

```{r}
library(tidyverse)
traits <- read_csv("https://jeffreyrstevens.github.io/dpavir2025/data/dog_breed_traits_clean.csv", show_col_types = FALSE)
```

2. Create `traits2` which adds a _coat_ column that combines _coat_type_ and _coat_length_ into single column delimited by "-".

```{r}
traits2 <- traits |> 
  unite("coat", contains("coat_"), sep = "-")
```

3. Split the _coat_ column into _type_ and _length_ and keep the original _coat_ column.

```{r}
traits2 |> 
  separate(coat, into = c("type", "length"), remove = FALSE)
```

4. Create `traits3` from `traits` that (1) removes the coat columns, (2) turns the ratings columns into long format, and (3) removes the _children_ row for _Bulldogs_.

```{r}
traits3 <- traits |> 
  select(-contains("coat_")) |> 
  pivot_longer(affectionate:grooming, names_to = "scale", values_to = "rating") |> 
  filter(breed != "Bulldogs" | scale != "children")
```

5. Create `traits4` from `traits3` that ensures a complete data set with all five ratings for all breeds (and fills in missing combinations with `NA`) and check for the missing Bulldog children row.

```{r}
traits4 <- traits3 |> 
  complete(breed, scale)
```

6. How could we copy the rating from the previous row into the Bulldog children row to replace the `NA`? (Note this is not a good idea in this case!)

```{r}
traits4 |> 
  fill(rating) |> 
  filter(breed == "Bulldogs")
```

7. From `traits`, generate all combinations of coat type and length **observed in the data**, excluding `NA`.

```{r}
traits |> 
  expand(nesting(coat_type, coat_length)) |> 
  drop_na()
```

8. From `traits`, generate **all possible** combinations of coat type and length, excluding `NA`.

```{r}
traits |> 
  expand(coat_type, coat_length) |> 
  drop_na()
```

