---
title: "Merging columns"
author: "Jeffrey R. Stevens"
date: "2025-02-28"
output: html_document
execute:
  freeze: auto
---

For these exercises, we'll use the dog breed traits data set along with the data on breed popularity rankings. 

1. Load tidyverse, download and import [`dog_breed_traits_clean.csv`](https://jeffreyrstevens.github.io/dpavir2025/data/dog_breed_traits_clean.csv) to `traits`, and import [`dog_breed_ranks.csv`](https://jeffreyrstevens.github.io/dpavir2025/data/dog_breed_ranks.csv) to `ranks`. **Make sure to download both files from the website, as they have changed or are new.**

```{r}
# >
```

2. Which breeds differ between `traits` and `ranks`?

```{r}
# >
```

3. Merge `traits` and `ranks` (in that order) to produce a data frame that includes breeds shared by both data sets. How many rows are there?

```{r}
# >
```


4. Merge `traits` and `ranks` (in that order) to produce a data frame that includes all breeds included in either data set. How many rows are there?

```{r}
# >
```

5. Merge `traits` and `ranks` (in that order) to produce a data frame that includes only breeds included in `traits`. How many rows are there?

```{r}
# >
```

5. Merge `traits` and `ranks` (in that order) to produce a data frame that includes only breeds included in `ranks`. How many rows are there?

```{r}
# >
```

6. Make `table4a` and `table4b` tidy then join them to replicate `table1`.

```{r}
# >
```

