---
title: "Numbers"
author: "Jeffrey R. Stevens"
date: "2023-03-06"
output: html_document
execute:
  freeze: auto
---

For these exercises, we'll use the dog breed traits and dog breed popularity rankings data sets. 

1. Load tidyverse and import [`dog_breed_traits_clean.csv`](https://jeffreyrstevens.github.io/dpavir2025/data/dog_breed_traits_clean.csv) to `traits`.

```{r}
library(tidyverse)
traits <- read_csv(here::here("data/dog_breed_traits_clean.csv"), show_col_types = FALSE)
```


2. Create a column of per-row means over all rating columns called _mean_ratings_ and assign to `traits2`.

```{r}
traits2 <- traits |> 
  rowwise() |> 
  mutate(mean_ratings = mean(c(affectionate, children, other_dogs, shedding, grooming, na.rm = TRUE)))
```

3. Convert _mean_ratings_ to a proportion in a column called _pmean_ratings_ and add to `traits2`.

```{r}
traits2 <- traits2 |> 
  mutate(pmean_ratings = mean_ratings / 5)
```

4. Apply a natural log transformation to the _pmeans_ratings_ vector.

```{r}
log(traits2$pmean_ratings)
```

5. Round _pmean_ratings_ to two decimal places.

```{r}
round(traits2$pmean_ratings, digits = 2)
```

6. Convert _pmean_ratings_ to scientific notation.

```{r}
format(traits2$pmean_ratings, scientific = TRUE)
```

7. Sum up the total grooming ratings for each coat type.

```{r}
traits |> 
  count(coat_type, wt = grooming)
```


8. Add inline R code to the following sentence in R Markdown to say how many rows have `NA` for _grooming_:

We are missing grooming data for [insert inline R code] breeds.
