---
title: "Strings"
author: "Jeffrey R. Stevens"
date: "2023-03-08"
output: html_document
execute:
  freeze: auto
---

For these exercises, we'll use the dog breed traits data set. 

1. Load tidyverse and import [`dog_breed_traits_clean.csv`](https://jeffreyrstevens.quarto.pub/dpavir/data/dog_breed_traits_clean.csv) to `traits`.

```{r}
# >
```


2. Return the first ten letters of the alphabet in upper case.

```{r}
# >
```


3. Create this character string and assign it to `mystring`: The elephant said "Hello" then 'Bye!'  Then view how it would be printed.

```{r}
# >
```


4. Find how many characters are in `mystring`.

```{r}
# >
```


5. Create a vector of the first four characters of the _coat_length_ column from `traits`.

```{r}
# >
```


6. Convert the _breed_ column of `traits` to sentence case.

```{r}
# >
```


7. Create series of sentences using _breed_ and _coat_length_ that states "[insert breed name] have a [insert coat length] coat" that uses the proper cases.

```{r}
# >
```


8. In the `mtcars` data set, extract the first two digits of the _mpg_ variable and the last three digits of the car names and combine them into a single string.
```{r}
# >
```


9. Print the `fruit` data set, then capitalize all first word letters in the data set, then capitalize all words in the data set.
```{r}
# >
```
