---
title: "Factors"
author: "Jeffrey R. Stevens"
date: "2025-03-12"
output: html_document
execute:
  freeze: auto
---

For these exercises, we'll use the dog breed traits data set. 

1. Load tidyverse, import [`dog_breed_traits_clean.csv`](https://jeffreyrstevens.github.io/dpavir2025/data/dog_breed_traits_clean.csv) to `traits`.

```{r}
# >
```


2. Convert both _coat_type_ and _coat_length_ into factors using `across()` and save as `traits2`.

```{r}
# >
```


3. Check the levels for both columns, one using a pipe and one without using a pipe.

```{r}
# >
```

4. Reorder the levels for _coat_length_ to be Short, Medium, Long (reassigned to `traits2`) and then check the levels.

```{r}
# >
```

5. Reorder the levels for _coat_type_ to be in the order of the most to least frequent coat type and then check the levels.

```{r}
# >
```

6. Relabel _coat_length_ to be Stubby, Mid, and Lush rather than Short, Medium, and Long.

```{r}
# >
```

7. The new AKC standard subsumes Rough coats with Wiry coats and Silky with Wavy. Please update the _coat_type_ variable accordingly.

```{r}
# >
```

