---
title: "Functions"
author: "Jeffrey R. Stevens"
date: "2023-03-27"
output: html_document
execute:
  freeze: auto
---


1. Write a function called `mystring` that takes a vector as an argument and returns the first three characters from the string. Test it on `words[1:10]`.

```{r}
library(stringr)
mystring <- function(x) {
  str_sub(x, 1, 3)
}
mystring(words[1:10])
```

2. Add an argument to `mystring()` that allows the user to control how many of the first characters should be returned. Test it on `words[1:10]` with 5 characters.

```{r}
mystring <- function(x, chars) {
  str_sub(x, 1, chars)
}
mystring(words[1:10], chars = 5)
```

3. Set the default number of characters returned by `mystring()` to be 3 and test that the default works and that you can override the default.

```{r}
mystring <- function(x, chars = 3) {
  str_sub(x, 1, chars)
}
mystring(words[1:10])
mystring(words[1:10], chars = 5)
```

4. Add a step that checks whether the inputted vector is a character string. If it is, continue to return the truncated strings. If the vector is not a character string, use the `stop()` function to stop the computation and return a message to the console telling the user that the vector was not a character vector. Test your function with a character vector, a numeric vector, and a logical vector.

```{r}
mystring <- function(x, chars = 3) {
  if (!is.character(x)) {
    stop("Please enter a character vector.")
  } else {
  str_sub(x, 1, chars)
  }
}

mystring(words[1:10])
# mystring(1:10)
# mystring(c(TRUE, FALSE))
```

5. Create a function called `parse_my_vector` that does the following:
  * Allows users to input a vector and a response to the argument `type` that determines whether the vector is a numeric (`"num"`), character (`"char"`), or logical (`"logical"`) vector. There should be no default value. If the user response does not match any of these three strings, stop with a message asking the user to specify one of the three strings.
  * For each type, checks whether the vector is actually the type specified by the user and stops with a message if they do not match.
  * For numeric vectors, multiplies by 10. For character vectors, extracts the first three characters. For logical vectors, returns the number of TRUE responses.
  * Before returning output, prints a message thanking the user.
  * Returns the original vector and output of the functions described above.

```{r}
parse_my_vector <- function(x, type) {
  if (type == "num") {
    if (is.numeric(x)) {
      output <- x / 10
    } else {
      stop("Type response does not match vector type.")
    }
  } else if (type == "char") {
    if (is.character(x)) {
      output <- str_sub(x, 1, 3)
    } else {
      stop("Type response does not match vector type.")
    }
  } else if (type == "logical") {
    output <- sum(x)
  } else {
    stop("Please enter either 'num', 'char', or 'logical'.")
  }
  message("Thank you!")
  list(x,output)
}
```

6. Check the following with `parse_my_vector()`:
  * x = 1:10, type = "num"
  * x = 1:10, type = "char"
  * x = words[1:10], type = "num"
  * x = words[1:10], type = "char"
  * x = c(TRUE, FALSE, TRUE), type = "num"
  * x = c(TRUE, FALSE, TRUE), type = "logical"
  * x = 1:10, type = "nums"

```{r}
parse_my_vector(x = 1:10, type = "num")
# parse_my_vector(x = 1:10, type = "char")
# parse_my_vector(x = words[1:10], type = "num")
parse_my_vector(x = words[1:10], type = "char")
# parse_my_vector(x = c(TRUE, FALSE, TRUE), type = "num")
parse_my_vector(x = c(TRUE, FALSE, TRUE), type = "logical")
# parse_my_vector(x = 1:10, type = "nums")
```

