---
title: "Iteration"
author: "Jeffrey R. Stevens"
date: "2023-03-29"
output: html_document
execute:
  freeze: auto
---


1. Write a for loop that calculates the mean bill length for each species in the `penguins` data set (don't use `group_by()`) and saves them as `species_means`.

```{r}
library(tidyverse)
library(palmerpenguins)
penguin_species <- unique(penguins$species)
species_means <- NA
for (i in penguin_species) {
  species_means[i] <- penguins |> 
    filter(species == i) |> 
    summarise(mean_bill_length = mean(bill_length_mm, na.rm = TRUE)) |> 
    pull()
}
species_means <- species_means[!is.na(species_means)]
```

2. Turn #1 into a function called `species_mean` that lets the user determine which variable to calculate the mean over.

```{r}
species_mean <- function(var) {
  penguin_species <- unique(penguins$species)
  species_means <- NA
  for (i in penguin_species) {
    species_means[i] <- penguins |> 
      filter(species == i) |> 
      summarise(mean({{var}}, na.rm = TRUE)) |> 
      pull()
  }
  species_means <- species_means[!is.na(species_means)]
  return(species_means)
}
species_mean(bill_length_mm)
```

3. Create a list `penguins_island` that separates the penguins data by island.

```{r}
penguins_island <- penguins |> 
  split(penguins$island)
```

4. Apply `map()` to find the number of observations for each year.

```{r}
map(penguins_island, nrow)
```


5. Apply `map()` to calculate the mean body weight for each island.

```{r}
map(penguins_island, ~ mean(.x$body_mass_g, na.rm = TRUE))
```

6. Rework #5 to return a numeric vector with values rounded to 1 decimal place.

```{r}
map_dbl(penguins_island, ~ round(mean(.x$body_mass_g, na.rm = TRUE), 1))
```

