---
title: "Grammar of graphics I"
author: "Jeffrey R. Stevens"
date: "2025-03-28"
output: html_document
execute:
  freeze: auto
---


1. Using the `mtcars` data, create a scatterplot of the fuel efficiency as a function of weight.

```{r}
library(tidyverse)
mtcars |> 
  ggplot(aes(x = wt, y = mpg)) +
  geom_point()
```

2. Repeat the plot but only with vehicles having 4 or 6 cylinders.

```{r}
mtcars |> 
  filter(cyl < 8) |> 
  ggplot(aes(x = wt, y = mpg)) +
  geom_point()
```

3. Repeat plot #1 but add a smooth line **underneath** the data points.

```{r}
mtcars |> 
  ggplot(aes(x = wt, y = mpg)) +
  geom_smooth() +
  geom_point()
```

4. Make a boxplot of fuel efficiency for each cylinder size.

```{r}
mtcars |> 
  mutate(cyl = as.factor(cyl)) |> 
  ggplot(aes(x = cyl, y = mpg)) +
  geom_boxplot()
```

5. Add " cylinders" to the end of each value in the cylinder column of data and replot #4.

```{r}
mtcars |> 
  mutate(cyl = paste0(cyl, " cylinders")) |> 
  ggplot(aes(x = cyl, y = mpg)) +
  geom_boxplot()
```

6. Replot #4 ordering the cylinders such that the median mpg increases from left to right.

```{r}
mtcars |> 
  mutate(cyl = as.factor(cyl)) |> 
  ggplot(aes(x = fct_reorder(cyl, mpg), y = mpg)) +
  geom_boxplot()
```
