---
title: "Color"
author: "Jeffrey R. Stevens"
date: "2023-04-07"
output: html_document
execute:
  freeze: auto
---


1. Using the `mtcars` data, create a scatterplot of the fuel efficiency as a function of weight with color based on the number of cylinders.

```{r}
library(tidyverse)
mtcars |> 
  mutate(cyl = as.factor(cyl)) |> 
  ggplot(aes(x = wt, y = mpg, color = cyl)) +
  geom_point()
```

2. Repeat the scatterplot but use `scale_color_brewer()` to set the palette to _Set1_.

```{r}
mtcars |> 
  mutate(cyl = as.factor(cyl)) |> 
  ggplot(aes(x = wt, y = mpg, color = cyl)) +
  geom_point() +
  scale_color_brewer(palette = "Set1")
```

3. Why did you use `scale_color_brewer()` not `scale_fill_brewer()` or `scale_color_distiller()`?


4. Repeat scatterplot #1 but use `scale_color_manual()` to set the three colors to _red_, _green_, and _blue_.

```{r}
mtcars |> 
  mutate(cyl = as.factor(cyl)) |> 
  ggplot(aes(x = wt, y = mpg, color = cyl)) +
  geom_point() +
  scale_color_manual(values = c("red", "green", "blue"))
```

5. Find three colors (either names or hex codes), and repeat scatterplot #4 with your own colors.

```{r}
mtcars |> 
  mutate(cyl = as.factor(cyl)) |> 
  ggplot(aes(x = wt, y = mpg, color = cyl)) +
  geom_point() +
  scale_color_manual(values = c("coral2", "palegreen4", "slateblue4"))
```


6. Using the `mtcars` data, create a scatterplot of the fuel efficiency as a function of weight with color based on horsepower using the _BuPu_ palette and reverse the direction of the gradient.

```{r}
mtcars |> 
  ggplot(aes(x = wt, y = mpg, color = hp)) +
  geom_point() +
  scale_color_distiller(palette = "BuPu", direction = 2)
```

7. Create a new column in `mtcars` that centers and scales displacement with the `scale()` function. Create a scatterplot of fuel efficiency as a function of weight with color based on the rescaled displacement, using a diverging scale of your choice.

```{r}
mtcars |> 
  mutate(disp2 = scale(disp)) |> 
  ggplot(aes(x = wt, y = mpg, color = disp2)) +
  geom_point() +
  scale_color_distiller(palette = "RdBu")
```

