---
title: "Adjusting axes"
author: "Jeffrey R. Stevens"
date: "2023-04-24"
output: html_document
execute:
  freeze: auto
---


1. Using the `mpg` data, create boxplots of highway fuel efficiency as a function of class.

```{r}
library(tidyverse)
mpg |> 
  ggplot(aes(x = class, y = hwy)) +
  geom_boxplot()
```


2. Zoom into the plot with y-axis limits of 15 and 40 without altering the data.

```{r}
mpg |> 
  ggplot(aes(x = class, y = hwy)) +
  geom_boxplot() +
  coord_cartesian(ylim = c(15, 40))
```


3. Change the y-axis limits to 15 and 40 but allow the statistical transformations to change the data.

```{r}
mpg |> 
  ggplot(aes(x = class, y = hwy)) +
  geom_boxplot() +
  ylim(15, 40)
```


4. Replot #1 but using a log10 scale.

```{r}
mpg |> 
  ggplot(aes(x = class, y = hwy)) +
  geom_boxplot() +
  scale_y_log10()
```


5. Replot #1 but with y-axis limits running from 0 to 50 and with labels in increments of 5 but no minor grid lines.

```{r}
mpg |> 
  ggplot(aes(x = class, y = hwy)) +
  geom_boxplot() +
  scale_y_continuous(limits = c(0, 50), breaks = seq(0, 50, 5)) +
  theme(panel.grid.minor = element_blank())
```


6. Replot #1 but create separate panels based on year (as rows) and cylinders (as columns) and allowing the scales to vary across rows.

```{r}
mpg |> 
  ggplot(aes(x = class, y = hwy)) +
  geom_boxplot() +
  facet_grid(rows = vars(year), cols = vars(cyl), scale = "free")
```


7. Replot #1 and assign it to `hwy_plot` and replot a similar version with city fuel efficiency named `cty_plot`. Then combine them into a compound plot labeled as subfigures A and B and save this figure on your computer as a PNG file.
```{r}
hwy_plot <- mpg |> 
  ggplot(aes(x = class, y = hwy)) +
  geom_boxplot()
cty_plot <- mpg |> 
  ggplot(aes(x = class, y = cty)) +
  geom_boxplot()
library(patchwork)
hwy_plot + cty_plot + plot_annotation(tag_levels = "A")
# ggsave("fuel_efficiency_class.png")
```

