---
title: "Dates and times"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(palmerpenguins)
options(tutorial.max.forked.procs = 10)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
```


## Dates

### Set-up

Working with dates requires the `{lubridate}` package, which is now part of the core tidyverse. Otherwise, you need to load it separately.

```{r echo = TRUE, message = FALSE}
library(tidyverse)
```

### Today's date

```{r todays-date}
question(
  'Which of the following commands return today\'s date (and only the date)?',
  answer('`Sys.Date()`', correct = TRUE),
  answer('`today()`', correct = TRUE),
  answer('`Sys.Time()`'),
  answer('`Sys.date()`'),
  answer('`Sys.time()`'),
  answer('`now()`'),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

## Creating dates and times

Remember the standard way to format dates is the ISO 8601 format: YYYY-MM-DD. 

**1. Type the code to convert the following dates or date/times to the standard format (e.g., 2021-06-11).**

**11 Jun 21**
```{r exercise4, exercise = TRUE}
mydate <- "11 Jun 21"
...
```
 
```{r exercise4-solution}
mydate <- "11 Jun 21"
dmy(mydate)
```
 
```{r exercise4-check}
grade_result(
  pass_if(~identical(.result, dmy(mydate)))
)
```

**2021-6-11**
```{r exercise5, exercise = TRUE}
mydate <- "2021-6-11"
...
```
 
```{r exercise5-solution}
mydate <- "2021-6-11"
ymd(mydate)
```
 
```{r exercise5-check}
grade_result(
  pass_if(~identical(.result, ymd(mydate)))
)
```

**June 11th 2021**
```{r exercise6, exercise = TRUE}
mydate <- "June 11th 2021"
...
```
 
```{r exercise6-solution}
mydate <- "June 11th 2021"
mdy(mydate)
```
 
```{r exercise6-check}
grade_result(
  pass_if(~identical(.result, mdy(mydate)))
)
```

**06/11/2021 15:00**
```{r exercise7, exercise = TRUE}
mydate <- "06/11/2021 15:00"
...
```
 
```{r exercise7-solution}
mydate <- "06/11/2021 15:00"
mdy_hm(mydate)
```
 
```{r exercise7-check}
grade_result(
  pass_if(~identical(.result, mdy_hm(mydate)))
)
```

## Extracting date/time components

Once your dates and times are formatted correctly, it is easy to extract individual components with the provided functions in `{lubridate}`.

**2. Find today's date and extract the full word for the month (no abbreviations) using a `{lubridate}` function.**
```{r exercise8, exercise = TRUE}
mydate <- today()
...
```
 
```{r exercise8-solution}
mydate <- today()
month(mydate, label = TRUE, abbr = FALSE)
```
 
```{r exercise8-check}
grade_result(
  pass_if(~identical(.result, month(mydate, label = TRUE, abbr = FALSE)))
)
```

###

**3. Create a character vector (not factors) of the full names of the _weekdays_ (not weekend days) using a `{lubridate}` function.**
```{r exercise10, exercise = TRUE}
...
```
 
```{r exercise10-solution}
as.character(wday(2:6, label = TRUE, abbr = FALSE))
```

```{r exercise10-check}
grade_result(
  pass_if(~identical(.result, as.character(wday(2:6, label = TRUE, abbr = FALSE))))
)
```

## Time spans

###

One of the great things about working with dates is that you can use math to calculate time spans.

**4. Calculate the number of days that have passed since the class started on January 22, 2025.**
```{r exercise11, exercise = TRUE}
...
```
 
```{r exercise11-solution}
today() - ymd("2025-01-22")
```
 
```{r exercise11-check}
grade_result(
  pass_if(~identical(.result, today() - ymd("2025-01-22")))
)
```

**5. Find the date 123 days from now.**
```{r exercise12, exercise = TRUE}
...
```
 
```{r exercise12-solution}
today() + 123
```
 
```{r exercise12-check}
grade_result(
  pass_if(~identical(.result, today() + 123))
)
```


### Wrap-up

Congratulations, you finished the tutorial!

To get credit for this assignment, replace my name with the first name that you submitted in the course introduction form in the code below and click *Run Code* to generate the text for you to submit to Canvas. 

```{r exit_function}
letter_nums <- letters[1:26]
generate_text <- function(x) {
  x <- tolower(x)
  offset <- 31
  if (!is.character(x)) stop("Oops! Did you surround your name with quotation marks?")
  if (x == "jeff") stop("Oops! You left my name in. Please type your name.")
  x <- stringr::str_sub(x, 1, 4)
  expanded <- stringr::str_split_fixed(x, "", 4)
  name_nums <- match(expanded, letter_nums)
  name_nums <- name_nums[!is.na(name_nums)]
  new_nums <- as.character(name_nums + offset)
  new_text <- paste("Now copy this string of numbers and submit them to Canvas:", stringr::str_c(new_nums, collapse = ""))
  return(new_text)
}
```

```{r exit_code, exercise = TRUE, exercise.setup = "exit_function"}
# replace my name below with your first name (surrounded by quotes)
first_name <- "Jeff"
generate_text(first_name)
```

### Assignment complete!

Great! Copy that code into Canvas, and you're all set for this tutorial.
