---
title: "Summarizing rows"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
options(tutorial.max.forked.procs = 10)
library(tidyr)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
```


## Data wrangling

### Setup

We're working with `{dplyr}` again, so let's load the `{dplyr}` package along with the `{palmerpenguins}` package to use the `penguins` data set.

```{r echo = TRUE, message = FALSE}
library(dplyr)
library(palmerpenguins)
head(penguins)
```

Now we're interested in aggregating data over rows.


## Summarizing data

### Overall summaries

While you can simply input data frame columns to functions to summarize them (e.g., `mean(penguins$bill_length_mm, na.rm = TRUE)`, you can also use `summarise()` to do this: `summarise(penguins, mean_bill = mean(bill_length_mm, na.rm = TRUE))`. One key advantage of using `summarise()` is that it can apply multiple functions and can apply to multiple columns.

**Calculate the mean and standard deviation of body mass (ignoring NAs) and label them as `mean` and `sd` respectively.**
```{r exercise8, exercise = TRUE}
summarise(penguins, ...)
```

```{r exercise8-check}
grade_result(
  pass_if(~identical(.result, summarise(penguins, mean = mean(body_mass_g, na.rm = TRUE), sd = sd(body_mass_g, na.rm = TRUE))))
)
```

###

**Calculate the mean of all of the bill and flipper measurements that include the text `_mm` in their column names (ignoring NAs).**

```{r exercise9, exercise = TRUE}
summarise(penguins, ...)
```

```{r exercise9-check}
grade_result(
  pass_if(~identical(.result, summarise(penguins, across(contains("_mm"), ~mean(.x, na.rm = TRUE)))))
)
```

## Summarizing groups

### 

The `summarise()` function is probably most useful when paired with the `group_by()` function, which allows separate summaries for each group or combination of groups. 

But since this involves applying multiple functions, before we do this, we need to review the pipe operator `|>`. The pipe allows us to chain multiple functions together to apply them to data in the specified order. Basically, a pipe takes the output from the previous command and submits it to the next command. Therefore, we typically use `group_by()` to group the data by columns then pipe the output to a `summarise()` statement to generate summaries per group.

**Calculate the mean and standard deviation of body size (labeled as `mean` and `sd` respectively) for male, female, and unknown sex penguins (ignoring NAs in body size).**
```{r exercise10, exercise = TRUE}
penguins |> 
  group_by(...) |> 
  summarise(...)
```

```{r exercise10-check}
grade_result(
  pass_if(~identical(.result, group_by(penguins, sex) %>% summarise(mean = mean(body_mass_g, na.rm = TRUE), sd = sd(body_mass_g, na.rm = TRUE))))
)
```

### 

**Calculate the mean body size (labeled as `mean`) for each species and sex (in that order) of penguin (ignoring NAs in body size).**

```{r exercise11, exercise = TRUE}
penguins |> 
  group_by(...) |> 
  summarise(...)
```

```{r exercise11-check}
grade_result(
  pass_if(~identical(.result, group_by(penguins, species, sex) %>% summarise(mean = mean(body_mass_g, na.rm = TRUE))))
)
```

## Slicing groups

### 

The `slice_()` functions allow you subset specific rows within each group. You can grab the first row(s) (`slice_head()`), last row(s) (`slice_tail()`), row(s) with smallest values in a column (`slice_min()`), row(s) with largest values in a column (`slice_max()`), or a random sample of rows  (`slice_sample()`). The `n` argument lets you select a specific number of rows per group, but the default is `n = 1`.

**Find the rows with the three subjects with the shortests bills for each year.**
```{r exercise12, exercise = TRUE}
penguins |> 
  group_by(...) |> 
  slice_...(...)
```

```{r exercise12-check}
grade_result(
  pass_if(~identical(.result, penguins |> group_by(year) |> slice_min(bill_length_mm, n = 3)))
)
```

### 

**Find the rows with each species' heaviest male and heaviest female (dropping `NA`s).**
```{r exercise13, exercise = TRUE}
penguins |> 
  drop_na(sex) |> 
  group_by(...) |> 
  slice_...(...)
```

```{r exercise13-check}
grade_result(
  pass_if(~identical(.result, penguins |> drop_na(sex) |> group_by(species, sex) |> slice_max(body_mass_g)))
)
```
```{r exercise13-solution}
penguins |> 
  drop_na(sex) |> 
  group_by(species, sex) |> 
  slice_max(body_mass_g)
```

```{r exercise13-code-check}
grade_code()
```

### 

### Wrap-up

Congratulations, you finished the tutorial!

To get credit for this assignment, replace my name with the first name that you submitted in the course introduction form in the code below and click *Run Code* to generate the text for you to submit to Canvas. 

```{r exit_function}
letter_nums <- letters[1:26]
generate_text <- function(x) {
  x <- tolower(x)
  offset <- 9
  if (!is.character(x)) stop("Oops! Did you surround your name with quotation marks?")
  if (x == "jeff") stop("Oops! You left my name in. Please type your name.")
  x <- stringr::str_sub(x, 1, 4)
  expanded <- stringr::str_split_fixed(x, "", 4)
  name_nums <- match(expanded, letter_nums)
  name_nums <- name_nums[!is.na(name_nums)]
  new_nums <- as.character(name_nums + offset)
  new_text <- paste("Now copy this string of numbers and submit them to Canvas:", stringr::str_c(new_nums, collapse = ""))
  return(new_text)
}
```

```{r exit_code, exercise = TRUE, exercise.setup = "exit_function"}
# replace my name below with your first name (surrounded by quotes)
first_name <- "Jeff"
generate_text(first_name)
```

### Assignment complete!

Great! Copy that code into Canvas, and you're all set for this tutorial.
