---
title: "Matching patterns"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
options(tutorial.max.forked.procs = 10)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
set.seed(3)
fruits <- fruit[1:20]
```


## Matching strings

### Regular expressions

`{stringr}` use _regular expressions_ to match characters in strings. This allows us to find, extract, and replace strings.

Here are a few metacharacters for regular expressions that help us generate pattern.s

| Metacharacters | Meaning |
|:--------------:|:--------|
|`.`| Wildcard--any character |
|`^`| Matches at beginning of string |
|`$`| Matches at end of string |
|`|`| Matches one pattern or another |
|`()`| Matches character group |
|`\d`| Matches numerical digits |
|`[]`| Matches any characters inside brackets |
|`[^]`| Matches any characters **not** inside brackets |
|`[a-z]`| Matches any lower case letters |
|`[A-Z]`| Matches any upper case letters |
|`[A-Za-z]`| Matches any letters |

###

To practice with regular expressions, we can use the `str_view_all()` function, which simply highlights patterns that match regular expressions passed to it. Let's work with the first 20 elements of the `fruit` data set, which we'll call `fruits`.

```{r}
fruits <- fruit[1:20]
```

Let's first view the letter "a" in `fruits`.

```{r echo = TRUE}
str_view_all(fruits, "a")
```

Strings that match the patterns are surrounded by `<>` just to show us where the pattern is matched. Now let's view letter "a"s that have any other characters in front of them.

```{r echo = TRUE}
str_view_all(fruits, ".a")
```

Notice apple and apricot are no longer marked in the view. Why not?

###

**Use `str_view_all()` to mark all _vowels_ in `fruits`.**
```{r exercise1, exercise = TRUE}
str_view_all(fruits, ...)
```
 
###

**Use `str_view_all()` to mark all _consonants_ in `fruits`.**
```{r exercise2, exercise = TRUE}
str_view_all(fruits, ...)
```
 
###

**Use `str_view_all()` to mark all fruits that end with "nut".**
```{r exercise3, exercise = TRUE}
str_view_all(fruits, ...)
```
 

## Detecting and extracting patterns

### 

We can detect whether a pattern was present in each element of a vector with `str_detect()`. Note this returns a logical vector, so we only get a TRUE/FALSE indication, but we can do our usual tricks with logical vectors.

**Use `str_detect()` to create a logical vector of fruits that include the pattern "berry".**
```{r exercise4, exercise = TRUE}
str_detect(fruits, ...)
```
 
```{r exercise4-check}
grade_result(
  pass_if(~identical(.result, str_detect(fruits, "berry")))
)
```


**Use `str_detect()` to determine how many fruits include the pattern "berry".**
```{r exercise5, exercise = TRUE}
...
```
 
```{r exercise5-check}
grade_result(
  pass_if(~identical(.result, sum(str_detect(fruits, "berry"))))
)
```

###

We can also pair `str_detect()` with `filter()` to return observations that match patterns in columns.

**Return the observations from the `starwars` data set for the names that include a "-".**
```{r exercise6, exercise = TRUE}
starwars |> 
  ...
```
 
```{r exercise6-check}
grade_result(
  pass_if(~identical(.result, starwars |> filter(str_detect(name, "-"))))
)
```

###

Often, instead of a logical vector detecting the presence of a pattern, we're interesting in extracting the elements of a vector that include a pattern. For this, we use `str_subset()`.

**Return the berries from `fruits`.**
```{r exercise7, exercise = TRUE}
str_subset(fruits, ...)
```
 
```{r exercise7-check}
grade_result(
  pass_if(~identical(.result, str_subset(fruits, "berry")))
)
```


**Return the `fruits` that have "black" or "blue" in them.**
```{r exercise8, exercise = TRUE}
...
```
 
```{r exercise8-check}
grade_result(
  pass_if(~identical(.result, str_subset(fruits, "blue|black")))
)
```


## Replacing patterns

### 

In addition to detecting and extracting patterns, we may want to replace them (e.g., to correct spelling mistakes). The `str_replace()` function replaces a pattern **the first time it appears within an element**. It will replace across all elements but only once per element. The `str_replace_all()` function will replace all instances across the whole vector.

**Make bell pepper and chili pepper one word (bellpepper and chilipepper) while leaving blood orange and canary melon as two words.**
```{r exercise9, exercise = TRUE}
str_replace(fruits, ...)
```
 
```{r exercise9-check}
grade_result(
  pass_if(~identical(.result, str_replace(fruits, " pepper", "pepper")))
)
```

###

**Replace just the first time "an" shows up in a string with "am".**
```{r exercise10, exercise = TRUE}
...(fruits, ...)
```
 
```{r exercise10-check}
grade_result(
  pass_if(~identical(.result, str_replace(fruits, "an", "am")))
)
```

**Replace any instance of "an" with "am".**
```{r exercise11, exercise = TRUE}
...(fruits, ...)
```
 
```{r exercise11-check}
grade_result(
  pass_if(~identical(.result, str_replace_all(fruits, "an", "am")))
)
```

### 

### Wrap-up

Congratulations, you finished the tutorial!

To get credit for this assignment, replace my name with the first name that you submitted in the course introduction form in the code below and click *Run Code* to generate the text for you to submit to Canvas. 

```{r exit_function}
letter_nums <- letters[1:26]
generate_text <- function(x) {
  x <- tolower(x)
  offset <- 14
  if (!is.character(x)) stop("Oops! Did you surround your name with quotation marks?")
  if (x == "jeff") stop("Oops! You left my name in. Please type your name.")
  x <- stringr::str_sub(x, 1, 4)
  expanded <- stringr::str_split_fixed(x, "", 4)
  name_nums <- match(expanded, letter_nums)
  name_nums <- name_nums[!is.na(name_nums)]
  new_nums <- as.character(name_nums + offset)
  new_text <- paste("Now copy this string of numbers and submit them to Canvas:", stringr::str_c(new_nums, collapse = ""))
  return(new_text)
}
```

```{r exit_code, exercise = TRUE, exercise.setup = "exit_function"}
# replace my name below with your first name (surrounded by quotes)
first_name <- "Jeff"
generate_text(first_name)
```

### Assignment complete!

Great! Copy that code into Canvas, and you're all set for this tutorial.
