---
title: "Numbers"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(palmerpenguins)
library(nycflights13)
options(tutorial.max.forked.procs = 10)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = TRUE)
```


## Floating points

###

Floating point numbers do not have a fixed number of digits after the decimal point. Because computer memory is limited, you cannot store numbers with infinite precision, so floating points are stored imprecisely. This results in weird situations where seemingly equivalent numbers are not equal to each other.

```{r}
(1 / 49) * 49 == 1
```

But this is because of the number of digits after the decimal.

```{r}
(1 / 50) * 50 == 1
```


### Comparing numbers

We can use several different functions and operators to compare numbers. The logical operator `==` requires pretty precise equivalence to return `TRUE`.

```{r}
1 == 1.000000000000001
1 == 1.0000000000000001
```

###

This illustrates the notion of _tolerance_ or what level of precision is tolerated in numerical comparisons.

We can also use the `all.equal()` base R function to compare numbers. If two numbers are the same it returns `TRUE`. If they are different, it returns the difference.

```{r}
all.equal(1, 1.0001)
```

With `all.equal()`, we can set the tolerance. For example, here we consider numbers the same if they are within 1/100 of each other.

```{r}
all.equal(1, 1.0001, tolerance = 0.01)
all.equal(1, 1.1, tolerance = 0.01)
```

###

There is also a `{dplyr}` function `near()` that compares vectors of floating point numbers with a built in tolerance.

```{r}
sqrt(2) ^ 2 == 2
near(sqrt(2) ^ 2, 2)
(vec1 <- c(sqrt(2) ^ 2, (3 ^ (1 / 3)) ^ 3))
(vec2 <- c(2, 3))
vec1 == vec2
near(vec1, vec2)
```


### Questions

**Which expression would return `TRUE`?**
```{r, message = FALSE}
x <- 1.01
y <- 1.011
```

```{r compare-number, echo=FALSE}
question(
  "",
  answer('`all.equal(x, y, tolerance = 0.001)`', correct = TRUE),
  answer('`x == y`'),
  answer('`dplyr::near(x, y)`'),
  answer('`all.equal(x, y)`'),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```


## Counting

###

As a reminder, we can use `count()` to count all of the rows for unique category levels. If you want to sum up the values in a column per group rather than just count the number of groups, add the `wt` argument with the column to be summed.


**1. Use `count()` to find the number of each species of penguins from the `penguins` dataset?**

```{r exercise1, exercise = TRUE}
penguins |> 
  count(...)
```
 
```{r exercise1-solution}
penguins |> 
  count(species)
```

```{r exercise1-check}
grade_result(
  pass_if(~identical(.result, penguins |> count(species)))
)
```

###

**2. What is the total mass of each species _per island_?**

```{r exercise2, exercise = TRUE}
penguins |> 
  count(...)
```
 
```{r exercise2-solution}
penguins |> 
  count(species, island, wt = body_mass_g)
```

```{r exercise2-check}
grade_result(
  pass_if(~identical(.result, penguins |> count(species, island, wt = body_mass_g)))
)
```

###

If we're trying to count inside a `summarise()`, we use the `n()` function---if, for example, you want to summarize means but also find the sample size. 

We can take this a step further to count the number of unique/distinct combinations in a group with `n_distinct()`. This is equivalent to `nrow(unique(...))`.

**3. Find the mean body mass and sample size per species.**

```{r exercise3, exercise = TRUE}
penguins |> 
  summarise(...)
```
 
```{r exercise3-solution}
penguins |> 
  summarise(.by = species,
            mean_mass = mean(body_mass_g, na.rm = TRUE),
            n())
```

```{r exercise3-check}
grade_result(
  pass_if(~identical(.result, penguins |> summarise(.by = species, mean_mass = mean(body_mass_g, na.rm = TRUE), n())))
)
```

**4. Find the number of species per island and label it `num_species`.**

```{r exercise4, exercise = TRUE}
penguins |> 
  summarise(...)
```
 
```{r exercise4-solution}
penguins |> 
  summarise(.by = island,
            num_species = n_distinct(species))
```

```{r exercise4-check}
grade_result(
  pass_if(~identical(.result, penguins |> summarise(.by = island, num_species = n_distinct(species))))
)
```

###

`NA`s can be counted by summing over the logical vector output by `is.na()`. Summing any logical vector will count up the number of `TRUE` elements. Relatedly, using `mean()` on a logical vector returns the proportion of `TRUE` elements.

**5. Find the number of individuals with no sex entered for each species and label it `missing_sex`.**

```{r exercise5, exercise = TRUE}
penguins |> 
  summarise(...)
```
 
```{r exercise5-solution}
penguins |> 
  summarise(.by = species,
            missing_sex = sum(is.na(sex)))
```

```{r exercise5-check}
grade_result(
  pass_if(~identical(.result, penguins |> summarise(.by = species, missing_sex = sum(is.na(sex)))
))
)
```


## Transforming numbers

### Rounding

When we calculate means, standard deviations, confidence intervals, etc., we often get values with many digits after the decimal. Usually, we only want one or two digits. To get this, we use `round()` and give the number digits after the decimal.

```{r}
round(1111.11111)
round(1111.11111, digits = 2)
round(1111.11111, digits = -2)
```

### Formatting

To control the formatting of numbers, we can use `format()`. This allows us to use scientific notation:

```{r}
format(0.0020, scientific = TRUE)
```

Or append extra zeros at the end:

```{r}
format(0.0020, nsmall = 4)
```

Note that the output of format is a _character string_. So it is no longer a numeric, and you can't do math on the output of `format()`.


**6. Set the argument `big.mark = ","` in the following format statement. What does that do?**

```{r exercise6, exercise = TRUE}
format(1000)
```
 
```{r exercise6-solution}
format(1000, big.mark = ",")
```


### Cutting numbers

To divide numbers into bins, use `cut()`. This creates a _factor_ with levels for each of the bins. You can set the bins with the `breaks` argument by passing a vector with the minimum value, intermediate break points, and maximum value.

**7. Use `cut()` to bin the `penguins` body mass data into the following bins: 0-3500, 3501-4800, and 4801-6300. The first 20 elements should look like this:**

```{r echo=FALSE}
cut(penguins$body_mass_g, breaks = c(0, 3500, 4800, 6300)) |> 
  head(20)
```



```{r exercise7, exercise = TRUE}
cut(...)
```
 
```{r exercise7-solution}
cut(penguins$body_mass_g, breaks = c(0, 3500, 4800, 6300))
```

```{r exercise7-check}
grade_result(
  pass_if(~identical(.result, cut(penguins$body_mass_g, breaks = c(0, 3500, 4800, 6300))))
)
```

###

**8. Now add labels for Low, Medium, and High for the three levels using the `labels` argument.**

```{r exercise8, exercise = TRUE}
cut(...)
```
 
```{r exercise8-solution}
cut(penguins$body_mass_g, breaks = c(0, 3500, 4800, 6300),
    labels = c("Low", "Medium", "High"))
```

```{r exercise8-check}
grade_result(
  pass_if(~identical(.result, cut(penguins$body_mass_g, breaks = c(0, 3500, 4800, 6300),
    labels = c("Low", "Medium", "High"))))
)
```



### Wrap-up

Congratulations, you finished the tutorial!

To get credit for this assignment, replace my name with the first name that you submitted in the course introduction form in the code below and click *Run Code* to generate the text for you to submit to Canvas. 

```{r exit_function}
letter_nums <- letters[1:26]
generate_text <- function(x) {
  x <- tolower(x)
  offset <- 20
  if (!is.character(x)) stop("Oops! Did you surround your name with quotation marks?")
  if (x == "jeff") stop("Oops! You left my name in. Please type your name.")
  x <- stringr::str_sub(x, 1, 4)
  expanded <- stringr::str_split_fixed(x, "", 4)
  name_nums <- match(expanded, letter_nums)
  name_nums <- name_nums[!is.na(name_nums)]
  new_nums <- as.character(name_nums + offset)
  new_text <- paste("Now copy this string of numbers and submit them to Canvas:", stringr::str_c(new_nums, collapse = ""))
  return(new_text)
}
```

```{r exit_code, exercise = TRUE, exercise.setup = "exit_function"}
# replace my name below with your first name (surrounded by quotes)
first_name <- "Jeff"
generate_text(first_name)
```

### Assignment complete!

Great! Copy that code into Canvas, and you're all set for this tutorial.
