---
title: "Iteration"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(palmerpenguins)
options(tutorial.max.forked.procs = 10)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
```



## For loops

### 

Because R is a vectorized programming language, most of the time, we can apply functions to vectors to solve our coding problems. But sometimes, you need to repeat commands a certain number of times. This is where `for` statements come in. They are especially useful for repeatedly sampling.

**Use a for loop to generate 20 random samples of 10 numbers from the sequence 1-100 and print them to the console. Use the `sample()` function to draw the samples, and feel free to look up how to use `sample()`.**

```{r exercise5, exercise = TRUE}
set.seed(1)
for (...) {
  print(sample(...))
}
```

```{r exercise5-solution}
set.seed(1)
for (i in 1:20) {
  print(sample(1:100, 10))
}
```

```{r exercise5-check}
set.seed(1)
grade_result(
  pass_if(~identical(.result, for (i in 1:20) {
  print(sample(1:100, 10))
}
  ))
)
```

## Mapping functions

###

Because R is vectorized, we can apply functions across vectors without having to use `for` loops. While base R uses `apply()` functions, tidyverse uses `map()` functions from the `{purrr}` library.

Different map functions output different data types.

```{r function-output}
quiz(
  question(
    'What kind of data type does `map()` return?',
    answer('List', correct = TRUE),
    answer('Data frame'),
    answer('Numeric vector'),
    answer('Character vector'),
    answer('Logical vector'),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question(
    'Which map function returns a data frame?',
    answer('`map_df()`', correct = TRUE),
    answer('`map()`'),
    answer('`map_dbl()`'),
    answer('`map_d()`'),
    answer('`map_data()`'),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question(
    'Check out the `purrr::list_cbind()` function documentation. What does this function do?',
    answer('`Combines list elements into a data frame`', correct = TRUE),
    answer('`Combines list elements into a vector`'),
    answer('`Flattens lists by removing one level of list hierarchy`'),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question(
    'Check out the `purrr::list_c()` function documentation. What does this function do?',
    answer('`Combines list elements into a vector`', correct = TRUE),
    answer('`Combines list elements into a data frame`'),
    answer('`Flattens lists by removing one level of list hierarchy`'),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```

###


**Use one of the `map()` functions to find the maximum value for each of the numeric columns in `penguins` (ignoring `NA`s) and return a numeric vector. You'll need to use `is.numeric` inside the `where()` function to select the numeric columns. See the `where()` documentation for more information.**

```{r exercise6, exercise = TRUE}
...
```

```{r exercise6-solution}
map_dbl(select(penguins, where(is.numeric)), ~ max(.x, na.rm = TRUE))
```

```{r exercise6-check}
set.seed(20210614)
grade_result(
  pass_if(~identical(.result, map_dbl(select(penguins, where(is.numeric)), ~ max(.x, na.rm = TRUE))
  ))
)
```

### 

### Wrap-up

Congratulations, you finished the tutorial!

To get credit for this assignment, replace my name with the first name that you submitted in the course introduction form in the code below and click *Run Code* to generate the text for you to submit to Canvas. 

```{r exit_function}
letter_nums <- letters[1:26]
generate_text <- function(x) {
  x <- tolower(x)
  offset <- 28
  if (!is.character(x)) stop("Oops! Did you surround your name with quotation marks?")
  if (x == "jeff") stop("Oops! You left my name in. Please type your name.")
  x <- stringr::str_sub(x, 1, 4)
  expanded <- stringr::str_split_fixed(x, "", 4)
  name_nums <- match(expanded, letter_nums)
  name_nums <- name_nums[!is.na(name_nums)]
  new_nums <- as.character(name_nums + offset)
  new_text <- paste("Now copy this string of numbers and submit them to Canvas:", stringr::str_c(new_nums, collapse = ""))
  return(new_text)
}
```

```{r exit_code, exercise = TRUE, exercise.setup = "exit_function"}
# replace my name below with your first name (surrounded by quotes)
first_name <- "Jeff"
generate_text(first_name)
```

### Assignment complete!

Great! Copy that code into Canvas, and you're all set for this tutorial.
