---
title: "Plotting distributions: boxplots and violin plots"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(palmerpenguins)
library(Hmisc)
options(tutorial.max.forked.procs = 10)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
```


## Introduction

This exercise covers plotting distributions including boxplots and violin plots.

### Set-up

Let's load `{tidyverse}` and `{palmerpenguins}`.

```{r echo = TRUE, message = FALSE}
library(tidyverse)
library(palmerpenguins)
```


## Violin plots

### 

Histograms and density plots can represent one or a few distributions, but comparing multiple distributions can be difficult with these plots. Violin plots provide a better visualization of multiple distributions. Violin plots are basically density plots that are rotated and mirrored to provide a symmetrical distribution of the data.

Violin plots use the `geom_violin()` geometric object, and here we create one for the body mass data for each species.

```{r violin, echo=TRUE, warning=FALSE}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_violin()
```

### Properties refresher

Just as a refresher, let's adjust both global and mapped properties of the plot.

```{r violin-color, warning=FALSE}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g, fill = species, color = species)) +
  geom_violin(alpha = 0.25, show.legend = FALSE)
```

**Map both violin area and borders to species, reduce transparency to 0.25, and remove the legend.**

```{r exercise1, exercise = TRUE, warning = FALSE}
penguins %>% 
  ggplot(...)
```

```{r exercise1-solution}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g, fill = species, color = species)) +
  geom_violin(alpha = 0.25, show.legend = FALSE)
```

### Adding other stats

Often it is useful to add other statistics to the plot instead of just violin plots.

```{r violin-statsummary, warning=FALSE}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_violin(fill = "lightblue", color = "steelblue", size = 1.25) +
  stat_summary(fun.data = mean_cl_normal, color = "steelblue")
```

**Add a point for the mean and error bars for 95% confidence intervals to each violin. Make the area color *lightblue*, the borders *steelblue* and line width 1.25, and points/error bars *steelblue* for all violins.**

```{r exercise6, exercise = TRUE, warning = FALSE}
penguins %>% 
  ggplot(...)
```

```{r exercise6-solution}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_violin(fill = "lightblue", color = "steelblue", linewidth = 1.25) +
  stat_summary(fun.data = mean_cl_normal, color = "steelblue")
```


## Boxplots

###

While histograms, density plots, and violin plots show fairly fine-grained representations, sometimes you want a courser representation. Boxplot provide these by showing the median, interquartile range, and other measures of range.

```{r boxplot, echo=TRUE, warning=FALSE}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_boxplot(fill = "lightblue", color = "steelblue")
```

The default behavior of boxplots is for the bar to represent the median, the box to represent the interquartile range (25th-75th percentile of the data), the whiskers to represent 1.5 times the interquartile range, and the points represent outliers (data outside of 1.5 times the interquartile range).

### Whisker length

You can change the whiskers to be the full range by specifying a very high value in `coef`, which is the value multiplied by the interquartile range to produce the whiskers (by default, `coef = 1.5`).

```{r boxplot-coef, echo=TRUE, warning=FALSE}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_boxplot(coef = 5)
```

### Outlier properties

To change properties of the outlier data points, you can prepend `outlier.` to the standard properties (e.g., `outlier.color`, `outlier.shape`).

```{r boxplot-outlier, warning=FALSE}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_boxplot(outlier.shape = 8, outlier.size = 3)
```

**Reproduce the outlier symbols and increase the size to 3. You may have to search for R symbols to find the symbol number.**

```{r exercise7, exercise = TRUE, warning = FALSE}
penguins %>% 
  ggplot(...)
```

```{r exercise7-solution}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_boxplot(outlier.shape = 8, outlier.size = 3)
```


### Notched boxplots

You can add notches to boxplots that represent roughly 95% confidence intervals around the median by setting `notch = TRUE`.

```{r boxplot-notched, warning=FALSE}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_boxplot(notch = TRUE)
```

**Set `notch = TRUE`.**

```{r exercise8, exercise = TRUE, warning = FALSE}
penguins %>% 
  ggplot(...)
```

```{r exercise8-solution}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_boxplot(notch = TRUE)
```

### Box width

Sometimes we want the boxes to be wider or narrow than the default. You can control box width with the `width` argument.

```{r box-width, echo=TRUE, warning=FALSE}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_boxplot(notch = TRUE, width = 0.5)
```

By default, the notch width is set to be 50% of the box width. You can change that with the `notchwidth` argument.

```{r notch-width, echo=TRUE, warning=FALSE}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_boxplot(notch = TRUE, width = 0.5, notchwidth = 0.75)
```

### Combining violin plots and boxplots

You may want to overlay boxplots over violin plots to illustrate the full distribution as well as summary statistics.


```{r violin-boxplot, warning=FALSE}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g, color = species, fill = species)) +
  geom_violin(alpha = 0.5, show.legend = FALSE) +
  geom_boxplot(alpha = 0.25, show.legend = FALSE)
```

**Overlay a boxplot on top of a violin plot for the penguin species' body masses. Color the lines and shaded areas based on the species but don't include a legend. Make the transparency 0.5 for the violin plot and 0.25 for the boxplot.**

```{r exercise9, exercise = TRUE, warning = FALSE}
penguins %>% 
  ggplot(...)
```

```{r exercise9-solution}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g, color = species, fill = species)) +
  geom_violin(alpha = 0.5, show.legend = FALSE) +
  geom_boxplot(alpha = 0.25, show.legend = FALSE)
```

### Wrap-up

Congratulations, you finished the tutorial!

To get credit for this assignment, replace my name with the first name that you submitted in the course introduction form in the code below and click *Run Code* to generate the text for you to submit to Canvas. 

```{r exit_function}
letter_nums <- letters[1:26]
generate_text <- function(x) {
  x <- tolower(x)
  offset <- 9
  if (!is.character(x)) stop("Oops! Did you surround your name with quotation marks?")
  if (x == "jeff") stop("Oops! You left my name in. Please type your name.")
  x <- stringr::str_sub(x, 1, 4)
  expanded <- stringr::str_split_fixed(x, "", 4)
  name_nums <- match(expanded, letter_nums)
  name_nums <- name_nums[!is.na(name_nums)]
  new_nums <- as.character(name_nums + offset)
  new_text <- paste("Now copy this string of numbers and submit them to Canvas:", stringr::str_c(new_nums, collapse = ""))
  return(new_text)
}
```

```{r exit_code, exercise = TRUE, exercise.setup = "exit_function"}
# replace my name below with your first name (surrounded by quotes)
first_name <- "Jeff"
generate_text(first_name)
```

### Assignment complete!

Great! Copy that code into Canvas, and you're all set for this tutorial.
