---
title: "Factors"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(palmerpenguins)
options(tutorial.max.forked.procs = 10)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
```


## Factors

###

As a reminder, the four core data types are character, double, integer, and logical. Factors _augmented integers_.


### Set-up
To work with factors, we'll use the `{forcats}` package, which is automatically loaded when you load the core tidyverse packages. We'll go ahead and load `{tidyverse}` since we'll need other tidyverse packages. We'll also load the `{palmerpenguins}` package to work with those data.

```{r echo = TRUE}
library(tidyverse)
library(palmerpenguins)
```

## Ordering factor levels

###

The factor levels for a variable may not be in the order that you want, for example to plot categorical variables. Therefore, we sometimes need to relevel a factor.

###

The `penguins` data frame has three variables that are factors: species, island, and sex.

```{r}
penguins
```

###

First, let's check the levels for the `species` variable.

```{r echo = TRUE}
levels(penguins$species)
```

They seem to be in alphabetical order. Let's say that we want the levels to be in the opposite order: Gentoo, Chinstrap, Adelie.

**Reverse the order of the levels by replacing the `...` with the proper `{forcats}` function. Note that since we are just working with the vector `penguins$species`, we don't need to mutate anything---we can just use one of the `{forcats}` functions to change the levels.**
```{r exercise1, exercise = TRUE}
penguins$species %>% 
  ... %>% 
  levels()
```
 
```{r exercise1-solution}
penguins$species %>% 
  fct_rev() %>% 
  levels()
```

```{r exercise1-check}
grade_result(
  pass_if(~identical(.result, penguins$species %>% 
  fct_rev() %>% 
  levels()))
)
```

###

**Now put the levels in the order Chinstrap, Gentoo, Adelie.**
```{r exercise1b, exercise = TRUE}
penguins$species %>% 
  ... %>% 
  levels()
```
 
```{r exercise1b-solution}
penguins$species %>% 
  fct_relevel("Chinstrap", "Gentoo", "Adelie") %>% 
  levels()
```

```{r exercise1b-check}
grade_result(
  pass_if(~identical(.result, penguins$species %>% 
  fct_relevel("Chinstrap", "Gentoo", "Adelie") %>% 
  levels()
  ))
)
```

## Recoding factors

###

We often need to change the factor level labels for a variable. This is termed _recoding_. We can also collapse or lump multiple levels into fewer levels.

###

Now we want the word "Island" to be added to each of the island names in the data frame. So we need to recode the existing factor levels.

**Recode the island variable to add "Island" to each factor level (e.g., "Biscoe Island", "Dream Island", and "Torgersen Island") using a `{forcats}` function (not a `{stringr}` function).**
```{r exercise2, exercise = TRUE}
penguins$island %>% 
  ... %>% 
  levels()
```
 
```{r exercise2-solution}
penguins$island %>% 
  fct_recode("Biscoe Island" = "Biscoe", "Dream Island" = "Dream", "Torgersen Island" = "Torgersen") %>% 
  levels()
```

```{r exercise2-check}
grade_result(
  pass_if(~identical(.result, penguins$island %>% 
  fct_recode("Biscoe Island" = "Biscoe", "Dream Island" = "Dream", "Torgersen Island" = "Torgersen") %>% 
  levels()))
)
```

###

Now we want to create a new column that provides each species' diet. Gentoo penguins are generalists (meaning they feed on a number of different food items) and Adelie and chinstrap penguins are specialists that feed primarily on krill. To do this, we will mutate the `species` variable into a new `diet` variable that recodes Gentoo to "generalist" and Adelie and Chinstrap to "specialist".

**Create a new column that recodes the species to their diets using the `fct_collapse()` function.**
```{r exercise3, exercise = TRUE}
penguins %>% 
  mutate(diet = fct_collapse(...))
```
 
```{r exercise3-solution}
penguins %>% 
  mutate(diet = fct_collapse(species, generalist = "Gentoo", specialist = c("Adelie", "Chinstrap")))
```

```{r exercise3-check}
grade_result(
  pass_if(~identical(.result, penguins %>% 
  mutate(diet = fct_collapse(species, generalist = "Gentoo", specialist = c("Adelie", "Chinstrap")))))
)
```

### 

### Wrap-up

Congratulations, you finished the tutorial!

To get credit for this assignment, replace my name with the first name that you submitted in the course introduction form in the code below and click *Run Code* to generate the text for you to submit to Canvas. 

```{r exit_function}
letter_nums <- letters[1:26]
generate_text <- function(x) {
  x <- tolower(x)
  offset <- 4
  if (!is.character(x)) stop("Oops! Did you surround your name with quotation marks?")
  if (x == "jeff") stop("Oops! You left my name in. Please type your name.")
  x <- stringr::str_sub(x, 1, 4)
  expanded <- stringr::str_split_fixed(x, "", 4)
  name_nums <- match(expanded, letter_nums)
  name_nums <- name_nums[!is.na(name_nums)]
  new_nums <- as.character(name_nums + offset)
  new_text <- paste("Now copy this string of numbers and submit them to Canvas:", stringr::str_c(new_nums, collapse = ""))
  return(new_text)
}
```

```{r exit_code, exercise = TRUE, exercise.setup = "exit_function"}
# replace my name below with your first name (surrounded by quotes)
first_name <- "Jeff"
generate_text(first_name)
```

### Assignment complete!

Great! Copy that code into Canvas, and you're all set for this tutorial.
