---
title: "Validating data"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
```



## Data validation

Validating that your data are as you expect is a critical step before processing them. Minimally, you should inspect your data sets with `str()` or `glimpse()` or look at them in the _Environment_ tab. You can build validation rules to check your data with assertions and testing.

## Summarizing data

There are many different ways to summarize data in R. `head()` is a way to see the first few rows of a data set using base R, whereas `glimpse()` is the tidyverse way. Try both `head()` and `glimpse()` on the `airquality` data set (built into base R).

```{r head-quiz}
quiz(
  question(
    "In what ways are the outputs of `head()` and `glimpse()` different for this data set?",
    answer("Only `glimpse()` gives the dimensions of the data", correct = TRUE),
    answer("`head()` organizes columns vertically, but `glimpse()` organizes them horizontally", correct = TRUE),
    answer("Only `glimpse()` gives the data types for columns", correct = TRUE),
    answer("Only `glimpse()` truncates the data"),
    answer("`glimpse()` only outputs the first six rows of data"),
  allow_retry = TRUE,
  random_answer_order = TRUE
  )
)
```

Now run the `summary()` function on the `airquality` data. 
```{r exercise1, exercise = TRUE}

```
 
```{r exercise1-solution}
summary(airquality)
```

```{r exercise1-code-check}
grade_code()
```


```{r summary-quiz}
quiz(
  question(
    "What is the median value for the `Solar.R` data column?",
    answer("205.0", correct = TRUE),
    answer("185.9"),
    answer("31.50"),
    answer("9.958"),
    answer("`NA`"),
  allow_retry = TRUE,
  random_answer_order = TRUE
  ),
  question(
    "How many `NA`s does the Wind column have?",
    answer("0", correct = TRUE),
    answer("37"),
    answer("7"),
    answer("20.7"),
    answer("NA"),
  allow_retry = TRUE,
  random_answer_order = TRUE
  ),
  question(
    "What does `NA` mean in R?",
    answer("Missing data", correct = TRUE),
    answer("Undefined numbers"),
    answer("Not a number"),
    answer("Numbers divided by 0"),
  allow_retry = TRUE,
  random_answer_order = TRUE
  )
)
```

## Validating with `{dataReporter}`

The [`{dataReporter}`](https://github.com/ekstroem/dataReporter) package provides a nice way of generating a data dictionary (they call them codebooks) while giving you a nice overview of your data. Install and load [`{dataReporter}`](https://github.com/ekstroem/dataReporter). Run `makeCodebook()` on `toyData` to explore this data set.


```{r validation-quiz}
quiz(
question(
  'How many rows does the `toyData` have?',
  answer('15', correct = TRUE),
  answer('6'),
  answer('7'),
  answer('9'),
  allow_retry = TRUE,
  random_answer_order = TRUE
),
question(
  'What data type is the `pill` column?',
  answer('factor', correct = TRUE),
  answer('numeric'),
  answer('character'),
  answer('logical'),
  allow_retry = TRUE,
  random_answer_order = TRUE
),
question(
  'How many unique values does the `events` column have?',
  answer('9', correct = TRUE),
  answer('15'),
  answer('7'),
  answer('1'),
  allow_retry = TRUE,
  random_answer_order = TRUE
),
question(
  'What percent of the `events` column is missing?',
  answer('20%', correct = TRUE),
  answer('13.33%'),
  answer('0%'),
  answer('100%'),
  allow_retry = TRUE,
  random_answer_order = TRUE
),
question(
  'What is the median value of the `change` column?',
  answer('0.33', correct = TRUE),
  answer('-0.62'),
  answer('0.66'),
  answer('1.6'),
  allow_retry = TRUE,
  random_answer_order = TRUE
),
question(
  'Which variable is described as a key with distinct values for each observation?',
  answer('`id`', correct = TRUE),
  answer('`region`'),
  answer('`events`'),
  answer('`pill`'),
  answer('`spotifysong`'),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
)
```

### 

### Wrap-up

Congratulations, you finished the tutorial!

To get credit for this assignment, replace my name with the first name that you submitted in the course introduction form in the code below and click *Run Code* to generate the text for you to submit to Canvas. 

```{r exit_function}
letter_nums <- letters[1:26]
generate_text <- function(x) {
  x <- tolower(x)
  offset <- 11
  if (!is.character(x)) stop("Oops! Did you surround your name with quotation marks?")
  if (x == "jeff") stop("Oops! You left my name in. Please type your name.")
  x <- stringr::str_sub(x, 1, 4)
  expanded <- stringr::str_split_fixed(x, "", 4)
  name_nums <- match(expanded, letter_nums)
  name_nums <- name_nums[!is.na(name_nums)]
  new_nums <- as.character(name_nums + offset)
  new_text <- paste("Now copy this string of numbers and submit them to Canvas:", stringr::str_c(new_nums, collapse = ""))
  return(new_text)
}
```

```{r exit_code, exercise = TRUE, exercise.setup = "exit_function"}
# replace my name below with your first name (surrounded by quotes)
first_name <- "Jeff"
generate_text(first_name)
```

### Assignment complete!

Great! Copy that code into Canvas, and you're all set for this tutorial.
