---
title: "Filtering rows"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
```


## Data wrangling

### Setup

We're working with `{dplyr}` again, so let's load the `{dplyr}` package along with the `{palmerpenguins}` package to use the `penguins` data set.

```{r echo = TRUE, message = FALSE}
library(dplyr)
library(palmerpenguins)
head(penguins)
```

Now we'll start working with rows for this lesson.


## Filtering rows {data-allow-skip=TRUE}

While base R uses the `subset()` function, tidyverse uses `filter()` to pull out rows based on data values. For instance, to filter the gentoo species of penguins:
```{r echo = TRUE}
filter(penguins, species == "Gentoo")
```

You can use multiple criteria within a column or combining multiple columns using logical operators `>`, `>=`, `<`, `<=`, `!=`, `==`, `%in%`.

**Produce a data frame of penguins with body masses between and including 4500-5000 g. **
```{r exercise1, exercise = TRUE}
filter(penguins, ...)
```
 
```{r exercise1-check}
grade_result(
  pass_if(~identical(.result, filter(penguins, body_mass_g >= 4500 & body_mass_g <=5000)))
)
```


**Produce a data frame of penguins measured in 2008 or 2009. **
```{r exercise2, exercise = TRUE}
filter(penguins, ...)
```
 
```{r exercise2-check}
grade_result(
  pass_if(~identical(.result, filter(penguins, year == 2008 | year == 2009)))
)
```

**Produce a data frame of penguins measured on Biscoe or Dream using the `%in%` operator. **
```{r exercise3, exercise = TRUE}
filter(...)
```
 
```{r exercise3-check}
grade_result(
  pass_if(~identical(.result, filter(penguins, island %in% c("Biscoe", "Dream"))))
)
```

**Produce a data frame of penguins of unknown sex by including only NAs. **
```{r exercise4, exercise = TRUE}
filter(...)
```
 
```{r exercise4-check}
grade_result(
  pass_if(~identical(.result, filter(penguins, is.na(sex))))
)
```


**Produce a data frame of penguins of unknown sex by excluding those found in the vector `c("male", "female")`. Hint: This one is tricky, but remember, `!` means NOT. Feel free to check the solution if you get stuck.**
```{r exercise5, exercise = TRUE}

```
 
```{r exercise5-check}
grade_result(
  pass_if(~identical(.result, filter(penguins, !sex %in% c("male", "female"))))
)
```


## Sorting rows

While most computations don't depend on the order of rows, some do (e.g., ranks) and sometimes it's just nice to visualize a data frame sorted by some of the column values. The `arrange()` function does this and hierarchically sorts based on the order of column names passed as arguments.

**Sort the rows by sex then body mass.**
```{r exercise6, exercise = TRUE}
arrange(penguins, ...)
```
 
```{r exercise6-check}
grade_result(
  pass_if(~identical(.result, arrange(penguins, sex, body_mass_g)))
)
```

**Now do the same, but sort body mass from highest to lowest.**
```{r exercise7, exercise = TRUE}
arrange(penguins, ...)
```
 
```{r exercise7-check}
grade_result(
  pass_if(~identical(.result, arrange(penguins, sex, desc(body_mass_g))))
)
```

### 

### Wrap-up

Congratulations, you finished the tutorial!

To get credit for this assignment, replace my name with the first name that you submitted in the course introduction form in the code below and click *Run Code* to generate the text for you to submit to Canvas. 

```{r exit_function}
letter_nums <- letters[1:26]
generate_text <- function(x) {
  x <- tolower(x)
  offset <- 25
  if (!is.character(x)) stop("Oops! Did you surround your name with quotation marks?")
  if (x == "jeff") stop("Oops! You left my name in. Please type your name.")
  x <- stringr::str_sub(x, 1, 4)
  expanded <- stringr::str_split_fixed(x, "", 4)
  name_nums <- match(expanded, letter_nums)
  name_nums <- name_nums[!is.na(name_nums)]
  new_nums <- as.character(name_nums + offset)
  new_text <- paste("Now copy this string of numbers and submit them to Canvas:", stringr::str_c(new_nums, collapse = ""))
  return(new_text)
}
```

```{r exit_code, exercise = TRUE, exercise.setup = "exit_function"}
# replace my name below with your first name (surrounded by quotes)
first_name <- "Jeff"
generate_text(first_name)
```

### Assignment complete!

Great! Copy that code into Canvas, and you're all set for this tutorial.
