---
title: "Coding basics"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(devtools)
knitr::opts_chunk$set(echo = FALSE)
```

## Terminology

Understanding the terminology used when coding in R will help you communicate with other R users and ask better questions when seeking help.

Answer the following questions about this code:

`mydata <- readr::read_csv(file = "data/mydata.csv")`

```{r terminology}
quiz(
  question("Which is the function name?",
           answer("`read_csv`", correct = TRUE),
           answer("`readr`", message = "Words before `::` refer to the package name."),
           answer("`file`", message = "This is the argument name."),
           answer("`data/mydata.csv`", message = "This is the argument value."),
           answer("`mydata`", message = "This is the object that the data are assigned to."),
           allow_retry = TRUE,
           random_answer_order = TRUE
  ),
  question("Which is the argument value?",
           answer("`read_csv`", message = "This is the function value."),
           answer("`readr`", message = "Words before `::` refer to the package name."),
           answer("`file`", message = "This is the argument name."),
           answer("`data/mydata.csv`", correct = TRUE),
           answer("`mydata`", message = "This is the object that the data are assigned to."),
           allow_retry = TRUE,
           random_answer_order = TRUE
  ),
  question("Which is the package name?",
           answer("`read_csv`", message = "This is the function value."),
           answer("`readr`", correct = TRUE),
           answer("`file`", message = "This is the argument name."),
           answer("`data/mydata.csv`", message = "This is the argument value."),
           answer("`mydata`", message = "This is the object that the data are assigned to."),
           allow_retry = TRUE,
           random_answer_order = TRUE
  ),
  question("Which is the object?",
           answer("`read_csv`", message = "This is the function value."),
           answer("`mydata`", correct = TRUE),
           answer("`readr`", message = "Words before `::` refer to the package name."),
           answer("`file`", message = "This is the argument name."),
           answer("`data/mydata.csv`", message = "This is the argument value."),
           allow_retry = TRUE,
           random_answer_order = TRUE
  ),
  question("Which is the argument name?",
           answer("`read_csv`", message = "This is the function value."),
           answer("`readr`", message = "Words before `::` refer to the package name."),
           answer("`file`", correct = TRUE),
           answer("`data/mydata.csv`", message = "This is the argument value."),
           answer("`mydata`", message = "This is the object that the data are assigned to."),
           allow_retry = TRUE,
           random_answer_order = TRUE
  )
)
```


## Assignment

Most of your coding in R will involve assigning values or output from functions to an object.


```{r assignment}
quiz(
  question("Which is the preferred method to assign a value to an object?",
           answer("`x <- 9`", correct = TRUE),
           answer("`x = 9`", message = "This can be confused with argument definition."),
           answer("`x == 9`", message = "This is the syntax for a conditional equals."),
           answer("`9 -> x`", message = "This is backwards."),
           allow_retry = TRUE,
           random_answer_order = TRUE
  ),
  question("Which method is preferred for assigning the same value to multiple objects?",
           answer("`a <- b <- c <- NA`", correct = TRUE),
           answer("`a <- NA, b <- NA, c <- NA`"),
           answer("`a <- NA; b <- NA; c <- NA`"),
           answer("`NA -> c -> b -> a`"),
           allow_retry = TRUE,
           random_answer_order = TRUE
  )
)
```


## Coding style

R is a flexible language with different ways of coding the same thing. While coding *style* is a matter of preference, some techniques can make your code more readable (especially to your instructor!).

```{r coding-style}
question("Which uses the preferred coding style?",
         answer("`sum(mydata, na.rm = TRUE) / 5`", correct = TRUE),
         answer("`sum(mydata,na.rm=TRUE)/5`", message = "This needs more space."),
         answer("`sum(mydata, na.rm = T) / 5`", message = "Avoid abbreviating TRUE."),
         answer("`sum (mydata, na.rm = TRUE) / 5`", message = "Don't put a space after fucntions."),
         allow_retry = TRUE,
         random_answer_order = TRUE
)
```


## Naming things

How you name objects, columns, files, and directories is also a preference, but there are best practices that make these things more human readable, machine readable, searchable, and sortable.

### 

```{r naming}
quiz(
  question("Which is an example of snake case?",
           answer("`my_data`", correct = TRUE),
           answer("`MyData`", message = "This is upper camel case."),
           answer("`myData`", message = "This is camel case."),
           answer("`MY_DATA`", message = "This is screaming snake case."),
           allow_retry = TRUE,
           random_answer_order = TRUE
  ),
  question("Which is a preferred name for an object that contains the t-test results for data from experiment 1?",
           answer("`expt1_ttest`", correct = TRUE),
           answer("`e1t`", message = "Not descriptive enough."),
           answer("`the_ttest_used_on_my_experiment_1_data`", message = "Waaaaaay too descriptive."),
           answer("`ttestforexpt1`", message = "Too cramped. Needs to separate words."),
           answer("`Experiment1TTest`", message = "Avoid upper case."),
           allow_retry = TRUE,
           random_answer_order = TRUE
  ),
    question("Why should you avoid have both `foo.R` and `Foo.R` files in the same directory?",
           answer("Not all operating systems can tell the difference", correct = TRUE),
           answer("They have redundant content"),
           answer("No one should have two foos"),
           answer("What the heck is a foo?"),
           allow_retry = TRUE,
           random_answer_order = TRUE
  ),
    question("Which is in the ISO 8601 standard date format?",
           answer("2021-04-07", correct = TRUE),
           answer("04-07-2021"),
           answer("07-04-2021"),
           answer("07-Apr-2021"),
           answer("Apr. 7, 2021"),
           allow_retry = TRUE,
           random_answer_order = TRUE
  )
)
```


### Wrap-up

Congratulations, you finished the tutorial!

To get credit for this assignment, replace my name with the first name that you submitted in the course introduction form in the code below and click *Run Code* to generate the text for you to submit to Canvas. 

```{r exit_function}
letter_nums <- letters[1:26]
generate_text <- function(x) {
  x <- tolower(x)
  offset <- 7
  if (!is.character(x)) stop("Oops! Did you surround your name with quotation marks?")
  if (x == "jeff") stop("Oops! You left my name in. Please type your name.")
  x <- stringr::str_sub(x, 1, 4)
  expanded <- stringr::str_split_fixed(x, "", 4)
  name_nums <- match(expanded, letter_nums)
  name_nums <- name_nums[!is.na(name_nums)]
  new_nums <- as.character(name_nums + offset)
  new_text <- paste("Now copy this string of numbers and submit them to Canvas:", stringr::str_c(new_nums, collapse = ""))
  return(new_text)
}
```

```{r exit_code, exercise = TRUE, exercise.setup = "exit_function"}
# replace my name below with your first name (surrounded by quotes)
first_name <- "Jeff"
generate_text(first_name)
```

### Assignment complete

Copy that code into Canvas, and you're all set for this tutorial.
