---
title: "Data structures"
author: "Jeffrey R. Stevens"
date: "2025-02-05"
output: html_document
execute:
  freeze: auto
---


### Vectors

Create a vector called  `dog_names` with the values _Bella_, _Daisy_, and _Max_.
```{r}
# >
```

Create a vector called `sex` with the values _Female_, _Male_, and _Male_.
```{r}
# >
```

Use the index operator to print to console only _Daisy_ and _Max_ from `dog_names`.
```{r}
# >
```

Replace the _Daisy_ entry with _Luna_ and print `dog_names` to console.
```{r}
# >
```

Test whether the name _Sophie_ is contained in the vector `dog_names`.
```{r}
# >
```


### Lists

Copy/paste and run this code: `(mylist <- list(a = 1:4, b = c(4, 3, 8, 5), c = LETTERS[10:15], d = c("yes", "yes")))`

```{r}
# >
```

Check the data types for each list element individually.
```{r}
# >
```

Check the data types for each list element with one command.
```{r}
# >
```

Combine list elements `a` and `b` into a single vector.
```{r}
# >
```

### Data frames

Create a data frame called `mydf` with three columns: `x`, `y`, and `z` and five rows. For `x` assign any five numbers, for `y` assign any five character strings, and for `z` assign any five logical values.
```{r}
# >
```

Create a data frame called `dogs` that combines the `dog_names` and `sex` vectors and print to console.
```{r}
# >
```

Print to console just Luna's row.
```{r}
# >
```

Print to console the number of rows in `dogs`.
```{r}
# >
```

