---
title: "Validating data"
author: "Jeffrey R. Stevens"
date: "2025-02-10"
output: html_document
execute:
  freeze: auto
---

For these exercises, we'll use the `mtcars` data set build into base R.

1. What are the dimensions of `mtcars`?
```{r}
dim(mtcars)
```

2. In one line of code, view the data types for all of the columns in `mtcars`.
```{r}
str(mtcars)
```

3. What is the range of values for the `mpg` column?
```{r}
range(mtcars$mpg)
```

4. What are all of the possible values used in `gear`?
```{r}
unique(mtcars$gear)
```

5. Check whether the value 5 is found in the `carb` column.
```{r}
5 %in% mtcars$carb
```

6. Do any columns have missing values?
```{r}
summary(mtcars)
```

7. What is the 3rd quartile for `mpg`?
```{r}
summary(mtcars$mpg)
```

8. Check whether all horsepower (`hp`) values fall between 50 and 300. Which row numbers fall out of this range?
```{r}
which(mtcars$hp < 50)
which(mtcars$hp > 300)
```

9. Make a codebook for `mtcars`.
```{r eval = FALSE}
#dataReporter::makeCodebook(mtcars, replace = TRUE)
```

