---
title: "Selecting columns"
author: "Jeffrey R. Stevens"
date: "2025-02-12"
output: html_document
execute:
  freeze: auto
---

For these exercises, we'll use the `iris` data set build into base R.

1. View `iris` to see what it looks like.
```{r}
head(iris)
```

2. Return a data frame with only the sepal columns using _inclusion_.
```{r}
library(dplyr)
select(iris, Sepal.Length, Sepal.Width)
```

3. Return a data frame with only the sepal columns using a _helper function_.
```{r}
select(iris, contains("Sepal"))
```

4. Return a data frame with the sepal and petal columns using a _helper function_.
```{r}
select(iris, Sepal.Length:Petal.Width)
```

5. Return a data frame with the sepal and petal columns using _exclusion_.
```{r}
select(iris, -Species)
```

6. Move `Species` to be the first column using `select()` and a helper function.
```{r}
select(iris, Species, everything())
```

7. Move `Species` to be the first column using `relocate()`.
```{r}
relocate(iris, Species, .before = 1)
```

8. Rename `Species` to `species` using `select()` and a helper function.
```{r}
select(iris, everything(), species = Species)
```

9. Rename `Species` to `species` using `rename()`.
```{r}
rename(iris, species = Species)
```

