---
title: "Piping"
author: "Jeffrey R. Stevens"
date: "2025-02-17"
output: html_document
execute:
  freeze: auto
---

For these exercises, we'll use the dog breed traits data set. 

1. Create a pipeline to do all of the following:

* assign pipeline to `traits`
* import data from <https://jeffreyrstevens.github.io/dpavir2025/data/dog_breed_traits.csv>
* subset only the columns _Breed_ through _Coat Length_
* remove the _Drooling Level_ column


```{r}
library(tidyverse)
traits <- read_csv("https://jeffreyrstevens.quarto.pub/dpavir/data/dog_breed_traits.csv") |> 
  select(Breed:`Coat Length`) |> 
  select(!`Drooling Level`)
```

2. Rename the column names to `"breed", "affectionate", "children", "other_dogs", "shedding", "grooming", "coat_type", "coat_length"` using this vector (either at the end of the last pipeline or in a separate line).
```{r}
colnames(traits) <- c("breed", "affectionate", "children", "other_dogs", "shedding", "grooming", "coat_type", "coat_length")
```

3. In a new pipe, use `traits` to do the following. 

* assign to `traits2`
* rescale all of the ratings columns by subtracting 1 from all of the values
* create a new column called _coat_ that combines the _coat_type_ and _coat_length_ columns by pasting the values of those two columns separated by `-`
* create a new column called _shed_ that dichotomizes _shedding_ such that values of 3 and above are "A lot" and values below 3 are "Not much" and places the new column after _shedding_
* calculate the mean rating for the _children_ and _other_dogs_ columns in a column called `mean_rating` and place it after _other_dogs_
```{r}
traits2 <- traits |> 
  mutate(across(affectionate:grooming, ~ .x - 1)) |> 
  mutate(coat = paste(coat_type, coat_length, sep = "-")) |> 
  mutate(shed = ifelse(shedding > 2, "A lot", "Not much"), .after = "shedding") |> 
  rowwise() %>%
  mutate(mean_rating = mean(c(children, other_dogs)), .after = "other_dogs")
```

4. In a new pipe, use `traits2` to do the following. 

* assign to `coat_grooming`
* subset only the _grooming_ and _coat_type_ columns
* run a linear model (`lm`) using the formula `grooming ~ coat_type` (remember to use a placeholder for the data)
* apply the `summary()` function
* print the results to console
```{r}
(coat_grooming <- traits2 |> 
  select(grooming, coat_type) |> 
  lm(grooming ~ coat_type, data = _) |> 
   summary())
```

