---
title: "Filtering rows"
author: "Jeffrey R. Stevens"
date: "2023-02-20"
output: html_document
execute:
  freeze: auto
---

For these exercises, we'll use a new clean version of the dog breed traits data set. 

1. Import data from <https://jeffreyrstevens.quarto.pub/dpavir/data/dog_breed_traits_clean.csv> and assign to `traits`.

```{r}
library(tidyverse)
traits <- read_csv("https://jeffreyrstevens.quarto.pub/dpavir/data/dog_breed_traits_clean.csv")
```

2. View dogs only with short coats.

```{r}
filter(traits, coat_length == "Short")
```

3. View a data frame excluding dogs with short coats.

```{r}
filter(traits, coat_length != "Short")
```

4. View dogs with double or silky coats.

```{r}
filter(traits, coat_type %in% c("Double", "Silky"))
```

5. View dogs with double or silky coats and shedding ratings 3 or below.

```{r}
filter(traits, coat_type %in% c("Double", "Silky") & shedding <= 3)
```

6. View dogs with `NA` for _coat_type_.

```{r}
filter(traits, is.na(coat_type))
```

7. View dogs with `NA` for any column.

```{r}
filter(traits, if_any(everything(), is.na))
```

8. View dogs not missing any data.

```{r}
drop_na(traits)
```

9. View dogs sorted by breed name.

```{r}
arrange(traits, breed)
```

10. View dogs sorted by coat type then coat length then affectionate rating.

```{r}
arrange(traits, coat_type, coat_length, affectionate)
```

