---
title: "Merging rows"
author: "Jeffrey R. Stevens"
date: "2025-03-03"
output: html_document
execute:
  freeze: auto
---

For these exercises, we'll use the dog breed traits and dog breed popularity rankings data sets. 

1. Load tidyverse, import [`dog_breed_traits_clean.csv`](https://jeffreyrstevens.github.io/dpavir2025/data/dog_breed_traits_clean.csv) to `traits`, import [`dog_breed_ranks.csv`](https://jeffreyrstevens.github.io/dpavir2025/data/dog_breed_ranks.csv) to `ranks`, and import [`dog_breed_popularity.csv`](https://jeffreyrstevens.github.io/dpavir2025/data/dog_breed_popularity.csv) to `popularity`.

```{r}
library(tidyverse)
traits <- read_csv(here::here("data/dog_breed_traits_clean.csv"), show_col_types = FALSE)
ranks <- read_csv(here::here("data/dog_breed_ranks.csv"), show_col_types = FALSE)
popularity <- read_csv(here::here("data/dog_breed_popularity.csv"), show_col_types = FALSE)
```


2. First, set a random seed by using `set.seed(2)`. Then create a subset of `ranks` that is a random selection of 10% of the rows, sort by breed name, and assign to `ranks2`.

```{r}
set.seed(2)
ranks2 <- slice_sample(ranks, prop = 0.1) |> 
  arrange(breed)
```


3. Use a filtering join to return the subset of `traits` that _matches_ the breeds in `ranks2` and assign this to `traits2`.

```{r}
traits2 <- traits |> 
  semi_join(ranks2, by = "breed")
```


4. Use a filtering join to return the subset of `traits` that _excludes_ the breeds in `ranks2`.

```{r}
traits |> 
  anti_join(ranks2, by = "breed")
```


5. Now we want to filter `traits` based on breeds in `popularity`. Notice that the breeds column in `popularity` is called _Breed_. This is problematic because the breed column in `traits` is called _breed_ and names are case-sensitive. Use `join_by()` to filter `traits` by breeds in `popularity`. How many rows are there?

```{r}
traits |> 
  semi_join(popularity, by = join_by(breed == Breed))
```


6. Use `filter()` (not joins) to return the subset of `traits` that _excludes_ the breeds in `ranks2`.

```{r}
traits |> 
  filter(!breed %in% ranks2$breed)
```


7. Append `traits2` to the bottom of itself.

```{r}
bind_rows(traits2, traits2)
```


8. Append `traits2` to the right of itself.

```{r}
bind_cols(traits2, traits2)
```


9. Append `traits2` to the right of `ranks2`. 

```{r}
bind_cols(ranks2, traits2)
```


10. Why is this not a good idea? What would be a better way to achieve this?

```{r}
left_join(ranks2, traits2)
```

