---
title: "Factors"
author: "Jeffrey R. Stevens"
date: "2025-03-12"
output: html_document
execute:
  freeze: auto
---

For these exercises, we'll use the dog breed traits data set. 

1. Load tidyverse, import [`dog_breed_traits_clean.csv`](https://jeffreyrstevens.github.io/dpavir2025/data/dog_breed_traits_clean.csv) to `traits`.

```{r}
library(tidyverse)
traits <- read_csv(here::here("data/dog_breed_traits_clean.csv"), show_col_types = FALSE)
set.seed(12)
breeds <- sample(traits$breed)
```


2. Convert both _coat_type_ and _coat_length_ into factors using `across()` and save as `traits2`.

```{r}
traits2 <- traits |> 
  mutate(across(contains("coat"), factor))
```


3. Check the levels for both columns, one using a pipe and one without using a pipe.

```{r}
levels(traits2$coat_type)
traits2 |> 
  pull(coat_length) |> 
  levels()
```

4. Reorder the levels for _coat_length_ to be Short, Medium, Long (reassigned to `traits2`) and then check the levels.

```{r}
traits2 <- traits2 |> 
  mutate(coat_length = fct_relevel(coat_length, "Short", "Medium", "Long"))
levels(traits2$coat_length)
```

5. Reorder the levels for _coat_type_ to be in the order of the most to least frequent coat type and then check the levels.

```{r}
traits2 <- traits2 |> 
  mutate(coat_type = fct_infreq(coat_type))
levels(traits2$coat_type)
```

6. Relabel _coat_length_ to be Stubby, Mid, and Lush rather than Short, Medium, and Long.

```{r}
traits2 <- traits2 |> 
  mutate(coat_length = fct_recode(coat_length, "Stubby" = "Short",
                                  "Mid" = "Medium",
                                  "Lush" = "Long"))
levels(traits2$coat_length)
```

7. The new AKC standard subsumes Rough coats with Wiry coats and Silky with Wavy. Please update the _coat_type_ variable accordingly.

```{r}
traits2 <- traits2 |> 
  mutate(coat_type = fct_collapse(coat_type, Wiry = c("Rough", "Wiry"),
                                  Wavy = c("Silky", "Wavy")))
levels(traits2$coat_type)
```

