---
title: "Dates and times"
author: "Jeffrey R. Stevens"
date: "2023-03-22"
output: html_document
execute:
  freeze: true
---

For these exercises, we'll use the dates data set. 

```{r echo = FALSE, include = FALSE}
library(tidyverse)
nrows <- 20
data <- data.frame(id = 1:nrows, 
                   birth_date = format(as_date(sample(as_date("1950-01-01"):as_date("2005-01-31"), nrows, replace = TRUE)), "%m/%d/%Y"),
                   test1 = as_date(sample(as_date("2022-01-01"):as_date("2022-03-31"), nrows, replace = TRUE)),
                   test2 = as_date(sample(as_date("2022-06-01"):as_date("2022-09-30"), nrows, replace = TRUE))
                   )
write_csv(data, here::here("data/dates.csv"))
```


1. Load tidyverse, import [`dates.csv`](https://jeffreyrstevens.quarto.pub/dpavir/data/dates.csv) to `dates`, and view the data set.

```{r}
library(tidyverse)
dates <- read_csv(here::here("data/dates.csv"), show_col_types = FALSE)
```


2. Convert `birth_date` to a date object and resave `dates`.

```{r}
dates <- dates |> 
  mutate(birth_date = mdy(birth_date))
```


3. Create a column called _time1_ that converts _test1_ to datetime and change the time zone to "America/Chicago".

```{r}
dates |> 
  mutate(time1 = as_datetime(test1, tz = "America/Chicago"))
```


4. Calculate each participant's age in years at the time of test 1, rounded to 1 decimal place, stored in _age_ and resave `dates`.

```{r}
dates <- dates |> 
  mutate(age = round(as.numeric(test1 - birth_date) / 365.25, 1))
```


5. Calculate the number of days between test 1 and test 2 for each participant and label this column _test_diff_ (and resave `dates`).

```{r}
dates <- dates |> 
  mutate(test_diff = test2 - test1)
```


6. Create `dates2` that subsets the participants who were born after January 1, 1970.

```{r}
dates2 <- dates |> 
  filter(birth_date > "1970-01-01")
```


7. Create a column named _diff_text_ that writes the following sentence for each participant in `dates2`: "Participant [insert _id_] (age: [insert _age_]) had test 1 on [insert _test1_] and test 2 on [insert _test2_], which were [insert _test_diff_] days apart."

```{r}
dates2 |> 
  mutate(diff_text = stringr::str_glue("Participant {id} (age: {age}) had test 1 on {test1} and test 2 on {test2}, which were {test_diff} days apart."))
```

