---
title: "Iteration"
author: "Jeffrey R. Stevens"
date: "2023-03-29"
output: html_document
execute:
  freeze: auto
---


1. Write a for loop that calculates the mean bill length for each species in the `penguins` data set (don't use `group_by()`) and saves them as `species_means`.

```{r}
library(tidyverse)
library(palmerpenguins)
penguin_species <- unique(penguins$species)
species_means <- NA
for (i in penguin_species) {
  species_means[i] <- penguins |> 
    filter(species == i) |> 
    summarise(mean_bill_length = mean(bill_length_mm, na.rm = TRUE)) |> 
    pull()
}
species_means <- species_means[!is.na(species_means)]
```

<!-- 2. Turn #1 into a function called `species_mean` that calculates the mean for one specific species that the user can specify. -->

<!-- ```{r} -->
<!-- species_mean <- function(sp) { -->
<!--   penguin_species <- unique(penguins$species) -->
<!--   species_means <- NA -->
<!--   for (i in penguin_species) { -->
<!--     species_means[i] <- penguins |>  -->
<!--       filter(species == i) |>  -->
<!--       summarise(mean({{var}}, na.rm = TRUE)) |>  -->
<!--       pull() -->
<!--   } -->
<!--   species_means <- species_means[!is.na(species_means)] -->
<!--   return(species_means) -->
<!-- } -->
<!-- species_mean(bill_length_mm) -->
<!-- ``` -->

2. Create a list `penguins_island` that separates the penguins data by island.

```{r}
penguins_island <- penguins |> 
  split(penguins$island)
```

3. Apply `map()` to find the number of observations for each year.

```{r}
map(penguins_island, nrow)
```


4. Apply `map()` to calculate the mean body weight for each island.

```{r}
map(penguins_island, ~ mean(.x$body_mass_g, na.rm = TRUE))
```

5. Rework #5 to return a _numeric vector_ with values rounded to 1 decimal place.

```{r}
map_dbl(penguins_island, ~ round(mean(.x$body_mass_g, na.rm = TRUE), 1))
```

