---
title: "Grammar of graphics II"
author: "Jeffrey R. Stevens"
date: "2023-04-03"
output: html_document
execute:
  freeze: auto
---


1. Using the `mtcars` data, create a scatterplot of the fuel efficiency as a function of weight.

```{r}
library(tidyverse)
mtcars |> 
  ggplot(aes(x = wt, y = mpg)) +
  geom_point()
```

2. Repeat the scatterplot, but make the points _violet_ open squares of size 5.

```{r}
mtcars |> 
  ggplot(aes(x = wt, y = mpg)) +
  geom_point(color = "violet", shape = 0, size = 5)
```

3. Repeat the scatterplot but with separate colors for cylinder levels.

```{r}
mtcars |> 
  ggplot(aes(x = wt, y = mpg, color = cyl)) +
  geom_point()
```

4. Why does the legend look like that? Fix it so there are separate colors for cylinder levels.

```{r}
mtcars |> 
  ggplot(aes(x = wt, y = mpg, color = as.factor(cyl))) +
  geom_point()
```

5. Overlay separate regression lines for each cylinder level.

```{r}
mtcars |> 
  ggplot(aes(x = wt, y = mpg, color = as.factor(cyl))) +
  geom_point() +
  geom_smooth(method = "lm")
```

6. Overlay a single _firebrick_ regression line over the points with a _firebrick1_-colored confidence band.

```{r}
mtcars |> 
  ggplot(aes(x = wt, y = mpg)) +
  geom_point(aes(color = as.factor(cyl))) +
  geom_smooth(method = "lm", color = "firebrick", fill = "firebrick1")
```

7. Plot the mean and standard error of the mean of fuel efficiency for each level of cylinder and color them _sienna_.

```{r}
mtcars |> 
  ggplot(aes(x = as.factor(cyl), y = mpg)) +
  stat_summary(color = "sienna")
```

