---
title: "Design and themes"
author: "Jeffrey R. Stevens"
date: "2023-04-05"
output: html_document
execute:
  freeze: auto
---


1. Using the `mtcars` data, create a scatterplot of the fuel efficiency as a function of weight with color based on the number of cylinders.

```{r}
library(tidyverse)
mtcars |> 
  mutate(cyl = as.factor(cyl)) |> 
  ggplot(aes(x = wt, y = mpg, color = cyl)) +
  geom_point()
```

2. Repeat the scatterplot but with classic, bw, and dark themes.

```{r}
mtcars |> 
  mutate(cyl = as.factor(cyl)) |> 
  ggplot(aes(x = wt, y = mpg, color = cyl)) +
  geom_point() +
  theme_classic()
mtcars |> 
  mutate(cyl = as.factor(cyl)) |> 
  ggplot(aes(x = wt, y = mpg, color = cyl)) +
  geom_point() +
  theme_bw()
mtcars |> 
  mutate(cyl = as.factor(cyl)) |> 
  ggplot(aes(x = wt, y = mpg, color = cyl)) +
  geom_point() +
  theme_dark()
```

3. Repeat the scatterplot from #1 but with no minor grid lines and no legend.

```{r}
mtcars |> 
  mutate(cyl = as.factor(cyl)) |> 
  ggplot(aes(x = wt, y = mpg, color = cyl)) +
  geom_point() +
  theme(panel.grid.minor = element_blank(),
        legend.position = "none")
```

4. Repeat the scatterplot from #1 but no minor grid lines for the x-axis (keep them for the y-axis) and move the legend inside the plot area and remove the legend title.

```{r}
mtcars |> 
  mutate(cyl = as.factor(cyl)) |> 
  ggplot(aes(x = wt, y = mpg, color = cyl)) +
  geom_point() +
  theme(panel.grid.minor.x = element_blank(),
        legend.position = c(0.8, 0.8),
        legend.title = element_blank())
```

5. Repeat the scatterplot from #1 but make the major y-axis grid lines black dashed lines and change the text font to 12 point Times font.

```{r}
mtcars |> 
  mutate(cyl = as.factor(cyl)) |> 
  ggplot(aes(x = wt, y = mpg, color = cyl)) +
  geom_point() +
  theme(panel.grid.major.y = element_line(color = "black", linetype = 2),
        text = element_text(family = "Times", size = 12))
```

6. Create a version of scatterplot #1 that minimizes the data-ink ratio by reducing non-data-ink.

```{r}
mtcars |> 
  mutate(cyl = as.factor(cyl)) |> 
  ggplot(aes(x = wt, y = mpg, color = cyl)) +
  geom_point() +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        legend.title = element_blank(),
        legend.position = c(0.8, 0.8))
```

