---
title: "Plotting distributions: boxplots"
author: "Francine Goh"
date: "2023-04-12"
output: html_document
execute:
  freeze: auto
---


1. Using the `penguins` data, create a boxplot that shows penguin flipper length by island without outliers.

```{r}
library(tidyverse)
library(palmerpenguins)
penguins %>%
  ggplot(aes(x = island, y = flipper_length_mm)) +
  geom_boxplot(outlier.shape = NA)
```

2. Add the means and standard error for each boxplot.

```{r}
penguins %>%
  ggplot(aes(x = island, y = flipper_length_mm)) +
  geom_boxplot(outlier.shape = NA) +
  stat_summary()
```

3. Switch from standard errors to confidence intervals, increase the size of the point, and color the box shading _chocolate_.

```{r}
penguins %>%
  ggplot(aes(x = island, y = flipper_length_mm)) +
  geom_boxplot(outlier.shape = NA, fill = "chocolate") +
  stat_summary(fun.data = mean_cl_normal, size = 0.75)
```

4. Fill the boxplots with color separately for each island and remove the legend.

```{r}
penguins %>%
  ggplot(aes(x = island, y = flipper_length_mm, fill = island)) +
  geom_boxplot(outlier.shape = NA) +
  stat_summary(fun.data = mean_cl_normal, size = 0.75) +
  theme(legend.position = "none")
```


5. Create a boxplot to show how flipper length differs for each species by island.

```{r}
penguins %>%
  ggplot(aes(x = island, y = flipper_length_mm, fill = island)) +
  geom_boxplot(outlier.shape = NA) +
  stat_summary(fun.data = mean_cl_normal, size = 0.75) +
  facet_wrap(vars(species)) +
  theme(legend.position = "none")
```


6. Recreate the boxplot #5 as a violin plot with a white background.

```{r}
penguins %>%
  ggplot(aes(x = island, y = flipper_length_mm, fill = island)) +
  geom_violin(outlier.shape = NA) +
  stat_summary(fun.data = mean_cl_normal, size = 0.75) +
  facet_wrap(vars(species)) +
  theme(legend.position = "none",
        panel.background = element_rect(fill = "white"))
```
