---
title: "Plotting x-y data: associations"
author: "Jeffrey R. Stevens"
date: "2023-04-17"
output: html_document
execute:
  freeze: auto
---


1. Using the `mpg` data, create a scatterplot of the highway fuel efficiency and city fuel efficiency.

```{r}
# >
```

2. Now add a dashed reference line showing equivalent values for the two axes and set the aspect ratio to 1.

```{r}
# >
```

3. Looks like there is a possibility of overplotting. Turn this into a bubble chart with dot size scaling to the number of data points for each dot and make the dot colors _steelblue_.

```{r}
# >
```

4. Add rugs to scatterplot #1 and change to minimal theme.

```{r}
# >
```


5. From scatterplot #1, color the dots by class, move the legend to the top left corner of the plot, and add marginal density plots.

```{r}
# >
```

6. Create a data frame called `mpg_num` that only includes variables with numeric values using the `where()` function. Then remove the _year_ column.

```{r}
# >
```


7. Create correlation plots of the numeric variables in `mpg_num` in both base R and using {GGally}'s `ggpairs()` function.
```{r}
# >
```


8. Create a correlation matrix of `mpg_num` with the `cor()` function. Then use `ggcorrplot()` from the {ggcorrplot} package to make a heatmap correlation plot with just the upper triangle of the matrix and using circles to represent correlation coefficient magnitude.

```{r}
# >
```

