---
title: "Adjusting axes"
author: "Jeffrey R. Stevens"
date: "2023-04-24"
output: html_document
execute:
  freeze: auto
---


1. Using the `mpg` data, create boxplots of highway fuel efficiency as a function of class.

```{r}
# >
```


2. Zoom into the plot with y-axis limits of 15 and 40 without altering the data.

```{r}
# >
```


3. Change the y-axis limits to 15 and 40 but allow the statistical transformations to change the data.

```{r}
# >
```


4. Replot #1 but using a log10 scale.

```{r}
# >
```


5. Replot #1 but with y-axis limits running from 0 to 50 and with labels in increments of 5 but no minor grid lines.

```{r}
# >
```


6. Replot #1 but create separate panels based on year (as rows) and cylinders (as columns) and allowing the scales to vary across rows.

```{r}
# >
```


7. Replot #1 and assign it to `hwy_plot` and replot a similar version with city fuel efficiency named `cty_plot`. Then combine them into a compound plot labeled as subfigures A and B and save this figure on your computer as a PNG file.
```{r}
# >
```

